---
title: "Population genomics analysis in Rhodnius ecuadoriensis"
author: "Enrique Hernandez"
date: "17 August 2019"
output: html_document
---

## Background

The purpose of this tutorial is understanding the population dynamics of domicile/peridomicile and sylvatic populations of main Chagas disease vector, Rhodnius ecuadoriensis, in sourthern Ecuador.

We will carry out conventional population analyses to test different hypothesis of geneflow among these triatomine populations.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# read in R packages

library(vcfR)
library(adegenet)
library(hierfstat)
library(pegas)
library(ape)
library(phangorn)
library(poppr)     # Make sure poppr is loaded if you haven't done so already.
library(magrittr)
library(stats)
library(ade4)
library(phytools)
library(mmod)
library(raster)
library(MASS)
library(ecodist)
library(ggplot2)
library(reshape)
library(reshape2)
library(tidyverse)
library(Stack)
library(corMLPE)
library(diveRsity)
library(graph4lg)
library(PopGenome)
library(AICcmodavg)
library(GUSbase)
library(GUSLD)
library(ggpubr)
library(RVAideMemoire)
library(piecewiseSEM)

```

## Read in genotype informatiom

In this step, we will read in R the genotype information obtained from the STACKS pipeline. We obtained 2552 high quality variants among 272 Rhodnius ecuadoriensis samples distributed in 25 populations among loja.

In addition, we will add extra information in the genind object such as population assigment, coordinates and other categories which will be useful for DAPC, AMOVA or spatial analyses.


```{r }
# set working directory
setwd("C:/Users/quiqu/Dropbox/Recuadoriensis_Analysis/C02_KNNimp")
# linux cluster directory, close when not in cluster
#setwd("~/Dropbox/Recuadoriensis_Analysis/C02_KNNimp")
# Read  in vcf file
rhod.vcf <- read.vcfR("./snps.loja.KNNimpR3_pophabtime.vcf")

# Convert vcf to genind object
rhod.genind <- vcfR2genind(rhod.vcf, sep = "/")

# Add population map from sample information file.
rhod.info <- read.csv("./C02KNNimpR3_pophabtime_strata.csv")
popmap <- as.factor(as.vector(rhod.info$pop))
rhod.genind@pop <- popmap
rhod.genind

# Add the samples coordinates in UTM from information file.
xy <- cbind(rhod.info$X_point, rhod.info$Y_point)
rhod.genind@other$xy <- as.data.frame(xy)
rhod.genind

# Append information file as strata in the genind object

strata(rhod.genind) <- rhod.info
rhod.genind

```

## Obtain each populations basic stats using diveRsity


```{r}
rhod.genepop <- genind_to_genepop(rhod.genind)

rhod.genepop

write.table(rhod.genepop, file = "rhod.genepop.txt", quote=FALSE, row.names=FALSE, col.names=FALSE)

divBasic("./rhod.genepop.txt", outfile = "rhod.basic.stats", bootstraps = 1000)



# compare with PopGenReport

rhod.popgenreport <- allel.rich(rhod.genind)

# save in ./rhod.basic.stats.PopGenReport
write_csv(as.data.frame(rhod.popgenreport$all.richness), file = "rhod.pop.popgenreport.all.richness.csv")
write_csv(as.data.frame(rhod.popgenreport$sum.richness), file = "rhod.pop.popgenreport.sum.richness.csv")
write_csv(as.data.frame(rhod.popgenreport$mean.richness), file = "rhod.pop.popgenreport.mean.richness.csv")
write_csv(as.data.frame(rhod.popgenreport$alleles.sampled), file = "rhod.pop.popgenreport.alleles.sampled.csv")
write_csv(as.data.frame(rhod.popgenreport$pop.sizes), file = "rhod.pop.popgenreport.pop.sizes.csv")




```

## calculate collection sites with dom and syl populations using divRsity


```{r}


rhod.domsyl <- popsub(rhod.genind, blacklist=c("SF", "LM", "AH", "ND", "HG", "TC", "VC", "BM", "GA", "EX", "YS", "TR", "SS", "AZ", "NJ", "NT", "RT", "TM"))
# Change the population map with the approrpiate domicile and sylvatic category.
popmap.hab <- as.factor(as.vector(rhod.domsyl@strata$pophab))
rhod.domsyl@pop <- popmap.hab

rhod.domsyl.genepop <- genind_to_genepop(rhod.domsyl)

rhod.domsyl.genepop

write.table(rhod.domsyl.genepop, file = "rhod.domsyl.genepop.txt", quote=FALSE, row.names=FALSE, col.names=FALSE)

divBasic("./rhod.domsyl.genepop.txt", outfile = "rhod.domsyl.basic.stats", bootstraps = 1000)



# compare with PopGenReport

rhod.domsyl.popgenreport <- allel.rich(rhod.domsyl)

# save in ./rhod.basic.stats.PopGenReport
write_csv(as.data.frame(rhod.domsyl.popgenreport$all.richness), file = "rhod.domsyl.popgenreport.all.richness.csv")
write_csv(as.data.frame(rhod.domsyl.popgenreport$sum.richness), file = "rhod.domsyl.popgenreport.sum.richness.csv")
write_csv(as.data.frame(rhod.domsyl.popgenreport$mean.richness), file = "rhod.domsyl.popgenreport.mean.richness.csv")
write_csv(as.data.frame(rhod.domsyl.popgenreport$alleles.sampled), file = "rhod.domsyl.popgenreport.alleles.sampled.csv")
write_csv(as.data.frame(rhod.domsyl.popgenreport$pop.sizes), file = "rhod.domsyl.popgenreport.pop.sizes.csv")



##  Test for differences in Allelic richness from PopGenReport
## BR

BRdom <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,1]))
BRdom <- mutate(BRdom, population = "BRdom")
BRsyl <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,2]))
BRsyl <- mutate(BRsyl, population = "BRsyl")
BR_df <- as_tibble(rbind(BRdom, BRsyl))

perm.t.test(value ~ population, data = BR_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -8.8011, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
##sample estimates:
#mean in group BRdom mean in group BRsyl 
#           1.300098            1.387051

## CE

CEdom <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,3]))
CEdom <- mutate(CEdom, population = "CEdom")
CEsyl <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,4]))
CEsyl <- mutate(CEsyl, population = "CEsyl")
CE_df <- as_tibble(rbind(CEdom, CEsyl))

perm.t.test(value ~ population, data = CE_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -3.0018, p-value = 0.0028
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CEdom mean in group CEsyl 
#           1.282509            1.314571 


## CG

CGdom <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,5]))
CGdom <- mutate(CGdom, population = "CGdom")
CGsyl <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,6]))
CGsyl <- mutate(CGsyl, population = "CGsyl")
CG_df <- as_tibble(rbind(CGdom, CGsyl))

perm.t.test(value ~ population, data = CG_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -1.6648, p-value = 0.09159
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CGdom mean in group CGsyl 
#           1.360400            1.377492

## CQ

CQdom <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,7]))
CQdom <- mutate(CQdom, population = "CQdom")
CQsyl <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,8]))
CQsyl <- mutate(CQsyl, population = "CQsyl")
CQ_df <- as_tibble(rbind(CQdom, CQsyl))

perm.t.test(value ~ population, data = CQ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)


#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 6.6059, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CQdom mean in group CQsyl 
#           1.340938            1.270880

## GL

GLdom <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,9]))
GLdom <- mutate(GLdom, population = "GLdom")
GLsyl <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,10]))
GLsyl <- mutate(GLsyl, population = "GLsyl")
GL_df <- as_tibble(rbind(GLdom, GLsyl))

perm.t.test(value ~ population, data = GL_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -8.4838, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group GLdom mean in group GLsyl 
#           1.261907            1.351793 
           
           
## HY

HYdom <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,11]))
HYdom <- mutate(HYdom, population = "HYdom")
HYsyl <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,12]))
HYsyl <- mutate(HYsyl, population = "HYsyl")
HY_df <- as_tibble(rbind(HYdom, HYsyl))

perm.t.test(value ~ population, data = HY_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -4.6673, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group HYdom mean in group HYsyl 
#           1.206059            1.249873


## SJ

SJdom <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,13]))
SJdom <- mutate(SJdom, population = "SJdom")
SJsyl <- as_tibble(as.vector(rhod.domsyl.popgenreport$all.richness[,14]))
SJsyl <- mutate(SJsyl, population = "SJsyl")
SJ_df <- as_tibble(rbind(SJdom, SJsyl))

perm.t.test(value ~ population, data = SJ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 0.15441, p-value = 0.8603
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group SJdom mean in group SJsyl 
#           1.210760            1.209374 


```


##A) Investigate gene flow between domicile and sylvatic habitats.

Calculate alpha diversity stats for different group categories: population, habitat and time of collection within communities

1 Domicile and sylvatic samples of groups within the 7 communities with samples from both habitats




```{r}

# Remove communities without samples collected in both habitats (domicile and sylvatic).
rhod.domsyl <- popsub(rhod.genind, blacklist=c("SF", "LM", "AH", "ND", "HG", "TC", "VC", "BM", "GA", "EX", "YS", "TR", "SS", "AZ", "NJ", "NT", "RT", "TM"))
# Change the population map with the approrpiate domicile and sylvatic category.
popmap.hab <- as.factor(as.vector(rhod.domsyl@strata$pophab))
rhod.domsyl@pop <- popmap.hab

# Basic information about the data. e.g
rhod.sum <- summary(rhod.domsyl)

# calculate per locus observed and expected heterozygocity {hierfstat}
rhod.bs <- basic.stats(rhod.domsyl, diploid = TRUE, digits = 2)
# Calculate Alleic richness
rhod.ar<- allelic.richness(rhod.domsyl)
#$overall
#   Ho    Hs    Ht   Dst   Htp  Dstp   Fst  Fstp   Fis  Dest 
#  0.18  0.17  0.22  0.05  0.23  0.05  0.23  0.24 -0.02  0.07

wc(rhod.domsyl)
#$FST
#[1] 0.27404
#$FIS
#[1] -0.02670102

# Calculate average Observed heterozygosity per population
mean(rhod.bs$Ho[,1])
# Expected heterozygosity Hs
mean(rhod.bs$Hs[,1])
# Calculating mean Fis per population
mean(rhod.bs$Fis[,1], na.rm=T)
#mean allelic richness by population
mean(rhod.ar$Ar[,1])
# Calculate loci on HWE
temp.domsyl <- seppop(rhod.domsyl, pop=rhod.domsyl@pop)

br.dom <- hw.test(temp.domsyl$BRdom, B=0) # 0.9137931
br.syl <- hw.test(temp.domsyl$BRsyl, B=0) # 0.9098746

ce.dom <- hw.test(temp.domsyl$CEdom, B=0) # 0.9188871
ce.syl <- hw.test(temp.domsyl$CEsyl, B=0) # 1

cg.dom <- hw.test(temp.domsyl$CGdom, B=0) # 0.9251567
cg.syl <- hw.test(temp.domsyl$CGsyl, B=0) # 0.9224138

cq.dom <- hw.test(temp.domsyl$CQdom, B=0) # 0.8981191
cq.syl <- hw.test(temp.domsyl$CQsyl, B=0) # 1

gl.dom <- hw.test(temp.domsyl$GLdom, B=0) # 1
gl.syl <- hw.test(temp.domsyl$GLsyl, B=0) # 0.9286834

hy.dom <- hw.test(temp.domsyl$HYdom, B=0) # 0.9310345
hy.syl <- hw.test(temp.domsyl$HYsyl, B=0) # 1

sj.dom <- hw.test(temp.domsyl$SJdom, B=0)  # 0.9392633
sj.syl <- hw.test(temp.domsyl$SJsyl, B=0)  # 0.9451411

# % of locus in HWE proportions

length(which(br.dom[,3]>0.05))/length(br.dom[,3]) # 0.0.914



boot.ppfis(rhod.domsyl) 



###########################################
#####    randomisatio test between domestic and sylvatic populations.
####
######################
####  Observed heterosaigosity 
####
###################################

## BR

BRdom <- as_tibble(as.vector(rhod.bs$Ho[,1]))
BRdom <- mutate(BRdom, population = "BRdom")
BRsyl <- as_tibble(as.vector(rhod.bs$Ho[,2]))
BRsyl <- mutate(BRsyl, population = "BRsyl")

BR_df <- as_tibble(Stack(BRdom, BRsyl))

ggplot(BR_df, aes(x= value, fill=population))+
    ylab("Loci") + xlab("Observed heterozygosities - Ho") +
    geom_histogram(bins = 30, colour = "black") +
    geom_vline(data = filter(BR_df, population == "BRdom"), aes(xintercept = mean(value)),
               size = 1, linetype = "dashed", colour = "red") +
    geom_vline(data = filter(BR_df, population == "BRsyl"), aes(xintercept = mean(value)),
               size = 1, linetype = "dashed", colour = "blue") +
    theme_classic()

t.test(value ~ population, data = BR_df, alternative = "two.sided")
perm.t.test(value ~ population, data = BR_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -10.576, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group BRdom mean in group BRsyl 
#          0.1557092           0.2320337 

##############


# Initialize a list to store the simulated test-statistics
simulated_means <- list()

nreps = 10000 # 5000 iterations

for(i in 1:nreps){
    
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- BR_df 
    
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "BRdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "BRsyl") 
                             %>% pull(value))
    
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <- 0.1557092 - 0.2320337
# -0.0763245

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# [1] 0



###############
#### CE

CEdom <- as_tibble(as.vector(rhod.bs$Ho[,3]))
CEdom <- mutate(CEdom, population = "CEdom")
CEsyl <- as_tibble(as.vector(rhod.bs$Ho[,4]))
CEsyl <- mutate(CEsyl, population = "CEsyl")
CE_df <- as_tibble(Stack(CEdom, CEsyl))

t.test(value ~ population, data = CE_df, alternative = "two.sided")
#t = -4.9597, df = 4794.5, p-value = 7.303e-07
perm.t.test(value ~ population, data = CE_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#	Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -4.9597, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CEdom mean in group CEsyl 
#          0.1562108           0.1953370 

# Initialize a list to store the simulated test-statistics
simulated_means <- list()
nreps = 5000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CE_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CEdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CEsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <- 0.1562108 - 0.1953370

ggplot() +
    ylab("Count") + xlab("Simulated mean difference") +
    geom_histogram(aes(x = simulated_means), bins = 30, 
                   fill = "grey", alpha = 0.4, colour = "black") +
    geom_vline(xintercept = diff_means_obs, size = 1, 
               linetype = "dashed", colour = "black") + 
    theme_classic()


abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# 0


###############
#### CG

CGdom <- as_tibble(as.vector(rhod.bs$Ho[,5]))
CGdom <- mutate(CGdom, population = "CGdom")
CGsyl <- as_tibble(as.vector(rhod.bs$Ho[,6]))
CGsyl <- mutate(CGsyl, population = "CGsyl")
CG_df <- as_tibble(Stack(CGdom, CGsyl))

t.test(value ~ population, data = CG_df, alternative = "two.sided")
# t = -3.1668, df = 5037.6, p-value = 0.001551

perm.t.test(value ~ population, data = CG_df, alternative = "two.sided", paired = FALSE, nperm = 10000)
#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -3.1668, p-value = 0.0005999
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CGdom mean in group CGsyl 
#          0.2066614           0.2308973 


# Initialize a list to store the simulated test-statistics
simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CG_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CGdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CGsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <- 0.2066614 - 0.2308973

ggplot() +
    ylab("Count") + xlab("Simulated mean difference") +
    geom_histogram(aes(x = simulated_means), bins = 30, 
                   fill = "grey", alpha = 0.4, colour = "black") +
    geom_vline(xintercept = diff_means_obs, size = 1, 
               linetype = "dashed", colour = "black") + 
    theme_classic()


abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# 0.0022


####
# CQ

CQdom <- as_tibble(as.vector(rhod.bs$Ho[,7]))
CQdom <- mutate(CQdom, population = "CQdom")
CQsyl <- as_tibble(as.vector(rhod.bs$Ho[,8]))
CQsyl <- mutate(CQsyl, population = "CQsyl")
CQ_df <- as_tibble(Stack(CQdom, CQsyl))

t.test(value ~ population, data = CQ_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = CQ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 0.68653, p-value = 0.4828
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CQdom mean in group CQsyl 
#          0.1769671           0.1716301

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CQ_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CQdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CQsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <- 0.1769671 - 0.1716301



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#0.491




####
# GL

GLdom <- as_tibble(as.vector(rhod.bs$Ho[,9]))
GLdom <- mutate(GLdom, population = "GLdom")
GLsyl <- as_tibble(as.vector(rhod.bs$Ho[,10]))
GLsyl <- mutate(GLsyl, population = "GLsyl")
GL_df <- as_tibble(Stack(GLdom, GLsyl))

t.test(value ~ population, data = GL_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = GL_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -4.1832, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group GLdom mean in group GLsyl 
#          0.1649765           0.1983621 

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- GL_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "GLdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "GLsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <- 0.1649765  - 0.1983621



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#1e-04


####
# HY

HYdom <- as_tibble(as.vector(rhod.bs$Ho[,11]))
HYdom <- mutate(HYdom, population = "HYdom")
HYsyl <- as_tibble(as.vector(rhod.bs$Ho[,12]))
HYsyl <- mutate(HYsyl, population = "HYsyl")
HY_df <- as_tibble(Stack(HYdom, HYsyl))

t.test(value ~ population, data = HY_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = HY_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -6.6636, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group HYdom mean in group HYsyl 
#          0.1240361           0.1730055  

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- HY_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "HYdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "HYsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <- 0.1240361 - 0.1730055

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#[1] 0



####
# SJ

SJdom <- as_tibble(as.vector(rhod.bs$Ho[,13]))
SJdom <- mutate(SJdom, population = "SJdom")
SJsyl <- as_tibble(as.vector(rhod.bs$Ho[,14]))
SJsyl <- mutate(SJsyl, population = "SJsyl")
SJ_df <- as_tibble(Stack(SJdom, SJsyl))

t.test(value ~ population, data = SJ_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = SJ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 0.21222, p-value = 0.8279
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group SJdom mean in group SJsyl 
#          0.1416144           0.1400980  

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- SJ_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "SJdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "SJsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <-  0.1416144 - 0.1400980 

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# 0.835


######################
####  Expected heterosigosity 
####
###################################

## BR

BRdom <- as_tibble(as.vector(rhod.bs$Hs[,1]))
BRdom <- mutate(BRdom, population = "BRdom")
BRsyl <- as_tibble(as.vector(rhod.bs$Hs[,2]))
BRsyl <- mutate(BRsyl, population = "BRsyl")
BR_df <- as_tibble(Stack(BRdom, BRsyl))

ggplot(BR_df, aes(x= value, fill=population))+
    ylab("Loci") + xlab("Observed heterozygosities - Ho") +
    geom_histogram(bins = 30, colour = "black") +
    geom_vline(data = filter(BR_df, population == "BRdom"), aes(xintercept = mean(value)),
               size = 1, linetype = "dashed", colour = "red") +
    geom_vline(data = filter(BR_df, population == "BRsyl"), aes(xintercept = mean(value)),
               size = 1, linetype = "dashed", colour = "blue") +
    theme_classic()

t.test(value ~ population, data = BR_df, alternative = "two.sided")
perm.t.test(value ~ population, data = BR_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -7.7488, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group BRdom mean in group BRsyl 
#          0.1741889           0.2190987 

##############


# Initialize a list to store the simulated test-statistics
simulated_means <- list()

nreps = 10000 # 5000 iterations

for(i in 1:nreps){
    
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- BR_df 
    
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "BRdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "BRsyl") 
                             %>% pull(value))
    
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <- 0.1741889 - 0.2190987 
# -0.0763245

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# [1] 0



###############
#### CE

CEdom <- as_tibble(as.vector(rhod.bs$Hs[,3]))
CEdom <- mutate(CEdom, population = "CEdom")
CEsyl <- as_tibble(as.vector(rhod.bs$Hs[,4]))
CEsyl <- mutate(CEsyl, population = "CEsyl")
CE_df <- as_tibble(Stack(CEdom, CEsyl))

t.test(value ~ population, data = CE_df, alternative = "two.sided")
#t = -4.9597, df = 4794.5, p-value = 7.303e-07
perm.t.test(value ~ population, data = CE_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#	Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -7.232, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CEdom mean in group CEsyl 
#          0.1610541           0.2125784 

# Initialize a list to store the simulated test-statistics
simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CE_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CEdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CEsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <- 0.1610541 - 0.2125784



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#[1] 0


###############
#### CG

CGdom <- as_tibble(as.vector(rhod.bs$Hs[,5]))
CGdom <- mutate(CGdom, population = "CGdom")
CGsyl <- as_tibble(as.vector(rhod.bs$Hs[,6]))
CGsyl <- mutate(CGsyl, population = "CGsyl")
CG_df <- as_tibble(Stack(CGdom, CGsyl))

t.test(value ~ population, data = CG_df, alternative = "two.sided")


perm.t.test(value ~ population, data = CG_df, alternative = "two.sided", paired = FALSE, nperm = 10000)
#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -3.0187, p-value = 0.0026
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CGdom mean in group CGsyl 
#          0.2057837           0.2242594 


# Initialize a list to store the simulated test-statistics
simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CG_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CGdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CGsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <- 0.2057837 - 0.2242594


abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# 0.0019


####
# CQ

CQdom <- as_tibble(as.vector(rhod.bs$Hs[,7]))
CQdom <- mutate(CQdom, population = "CQdom")
CQsyl <- as_tibble(as.vector(rhod.bs$Hs[,8]))
CQsyl <- mutate(CQsyl, population = "CQsyl")
CQ_df <- as_tibble(Stack(CQdom, CQsyl))

t.test(value ~ population, data = CQ_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = CQ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 1.8986, p-value = 0.05579
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CQdom mean in group CQsyl 
#          0.1953174           0.1820141 

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CQ_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CQdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CQsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <-  0.1953174  - 0.1820141 



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#0.0599




####
# GL

GLdom <- as_tibble(as.vector(rhod.bs$Hs[,9]))
GLdom <- mutate(GLdom, population = "GLdom")
GLsyl <- as_tibble(as.vector(rhod.bs$Hs[,10]))
GLsyl <- mutate(GLsyl, population = "GLsyl")
GL_df <- as_tibble(Stack(GLdom, GLsyl))

t.test(value ~ population, data = GL_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = GL_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -8.2027, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group GLdom mean in group GLsyl 
#          0.1600157           0.2144788 

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- GL_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "GLdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "GLsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <-0.1600157 -  0.2144788



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#0


####
# HY

HYdom <- as_tibble(as.vector(rhod.bs$Hs[,11]))
HYdom <- mutate(HYdom, population = "HYdom")
HYsyl <- as_tibble(as.vector(rhod.bs$Hs[,12]))
HYsyl <- mutate(HYsyl, population = "HYsyl")
HY_df <- as_tibble(Stack(HYdom, HYsyl))

t.test(value ~ population, data = HY_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = HY_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -5.3656, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group HYdom mean in group HYsyl 
#          0.1164420           0.1457994 

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- HY_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "HYdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "HYsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <-  0.1164420    -       0.1457994

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#[1] 0



####
# SJ

SJdom <- as_tibble(as.vector(rhod.bs$Hs[,13]))
SJdom <- mutate(SJdom, population = "SJdom")
SJsyl <- as_tibble(as.vector(rhod.bs$Hs[,14]))
SJsyl <- mutate(SJsyl, population = "SJsyl")
SJ_df <- as_tibble(Stack(SJdom, SJsyl))

t.test(value ~ population, data = SJ_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = SJ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#10000 permutations
#t = 0.039822, p-value = 0.9845
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group SJdom mean in group SJsyl 
#          0.1182602           0.1180564 

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- SJ_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "SJdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "SJsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

diff_means_obs <-  0.1182602  -  0.1180564 

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# 0.9663



########################################################
####
####  Fis 
####
###################################

## BR

BRdom <- as_tibble(as.vector(rhod.bs$Fis[,1]))
BRdom <- mutate(BRdom, population = "BRdom")
BRsyl <- as_tibble(as.vector(rhod.bs$Fis[,2]))
BRsyl <- mutate(BRsyl, population = "BRsyl")
BR_df <- as_tibble(Stack(BRdom, BRsyl))

ggplot(BR_df, aes(x= value, fill=population))+
    ylab("Loci") + xlab("Observed heterozygosities - Ho") +
    geom_histogram(bins = 30, colour = "black") +
    geom_vline(data = filter(BR_df, population == "BRdom"), aes(xintercept = mean(value)),
               size = 1, linetype = "dashed", colour = "red") +
    geom_vline(data = filter(BR_df, population == "BRsyl"), aes(xintercept = mean(value)),
               size = 1, linetype = "dashed", colour = "blue") +
    theme_classic()

t.test(value ~ population, data = BR_df, alternative = "two.sided")
perm.t.test(value ~ population, data = BR_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 7.4276, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group BRdom mean in group BRsyl 
#         0.09788364         -0.03373479 

##############


# Initialize a list to store the simulated test-statistics
simulated_means <- list()

nreps = 10000 # 5000 iterations

for(i in 1:nreps){
    
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- BR_df 
    
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "BRdom") 
                           %>% pull(value), na.rm=T)
    mean_syl_sim <- mean(reshuffled %>% filter(population == "BRsyl") 
                             %>% pull(value), na.rm=T)
    
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)


mean_BRdom <- mean(BR_df %>% filter(population == "BRdom") 
                           %>% pull(value), na.rm=T)
mean_BRsyl <- mean(BR_df %>% filter(population == "BRsyl") 
                             %>% pull(value), na.rm=T)


diff_means_obs <- mean_BRdom -  mean_BRsyl


abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# [1]0




###############
#### CE

CEdom <- as_tibble(as.vector(rhod.bs$Fis[,3]))
CEdom <- mutate(CEdom, population = "CEdom")
CEsyl <- as_tibble(as.vector(rhod.bs$Fis[,4]))
CEsyl <- mutate(CEsyl, population = "CEsyl")
CE_df <- as_tibble(Stack(CEdom, CEsyl))

t.test(value ~ population, data = CE_df, alternative = "two.sided")
#t = -4.9597, df = 4794.5, p-value = 7.303e-07
perm.t.test(value ~ population, data = CE_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#	Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 4.7211, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CEdom mean in group CEsyl 
#         0.04642217         -0.06176784 

# Initialize a list to store the simulated test-statistics
simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CE_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CEdom") 
                           %>% pull(value), na.rm=T)
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CEsyl") 
                             %>% pull(value), na.rm=T)
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

mean_CEdom <- mean(CE_df %>% filter(population == "CEdom") 
                           %>% pull(value), na.rm=T)
mean_CEsyl <- mean(CE_df %>% filter(population == "CEsyl") 
                             %>% pull(value), na.rm=T)



diff_means_obs <- mean_CEdom - mean_CEsyl



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#[1] 0


###############
#### CG

CGdom <- as_tibble(as.vector(rhod.bs$Fis[,5]))
CGdom <- mutate(CGdom, population = "CGdom")
CGsyl <- as_tibble(as.vector(rhod.bs$Fis[,6]))
CGsyl <- mutate(CGsyl, population = "CGsyl")
CG_df <- as_tibble(Stack(CGdom, CGsyl))

t.test(value ~ population, data = CG_df, alternative = "two.sided")


perm.t.test(value ~ population, data = CG_df, alternative = "two.sided", paired = FALSE, nperm = 10000)
#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 3.4811, p-value = 4e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CGdom mean in group CGsyl 
#        0.006835616        -0.052172662 


# Initialize a list to store the simulated test-statistics
simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CG_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CGdom") 
                           %>% pull(value), na.rm=T)
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CGsyl") 
                             %>% pull(value), na.rm=T)
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

mean_CEdom <- mean(CE_df %>% filter(population == "CEdom") 
                           %>% pull(value), na.rm=T)
mean_CEsyl <- mean(CE_df %>% filter(population == "CEsyl") 
                             %>% pull(value), na.rm=T)



diff_means_obs <- mean_CEdom - mean_CEsyl



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# 0.0019


####
# CQ

CQdom <- as_tibble(as.vector(rhod.bs$Fis[,7]))
CQdom <- mutate(CQdom, population = "CQdom")
CQsyl <- as_tibble(as.vector(rhod.bs$Fis[,8]))
CQsyl <- mutate(CQsyl, population = "CQsyl")
CQ_df <- as_tibble(Stack(CQdom, CQsyl))

t.test(value ~ population, data = CQ_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = CQ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 7.4273, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CQdom mean in group CQsyl 
#         0.08188276         -0.09090909  

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CQ_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CQdom") 
                           %>% pull(value), na.rm=T)
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CQsyl") 
                             %>% pull(value), na.rm=T)
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)


mean_CQdom <- mean(CQ_df %>% filter(population == "CQdom") 
                           %>% pull(value), na.rm=T)
mean_CQsyl <- mean(CQ_df %>% filter(population == "CQsyl") 
                             %>% pull(value), na.rm=T)

diff_means_obs <-  mean_CQdom  - mean_CQsyl 



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#0


####
# GL

GLdom <- as_tibble(as.vector(rhod.bs$Fis[,9]))
GLdom <- mutate(GLdom, population = "GLdom")
GLsyl <- as_tibble(as.vector(rhod.bs$Fis[,10]))
GLsyl <- mutate(GLsyl, population = "GLsyl")
GL_df <- as_tibble(Stack(GLdom, GLsyl))

t.test(value ~ population, data = GL_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = GL_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -4.3332, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group GLdom mean in group GLsyl 
#        -0.08298266          0.02215873 

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- GL_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "GLdom") 
                           %>% pull(value), na.rm=T)
    mean_syl_sim <- mean(reshuffled %>% filter(population == "GLsyl") 
                             %>% pull(value), na.rm=T)
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

mean_GLdom <- mean(GL_df %>% filter(population == "GLdom") 
                           %>% pull(value), na.rm=T)
mean_GLsyl <- mean(GL_df %>% filter(population == "GLsyl") 
                             %>% pull(value), na.rm=T)

diff_means_obs <- -0.1051414



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#1


####
# HY

HYdom <- as_tibble(as.vector(rhod.bs$Fis[,11]))
HYdom <- mutate(HYdom, population = "HYdom")
HYsyl <- as_tibble(as.vector(rhod.bs$Fis[,12]))
HYsyl <- mutate(HYsyl, population = "HYsyl")
HY_df <- as_tibble(Stack(HYdom, HYsyl))

t.test(value ~ population, data = HY_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = HY_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 7.3065, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group HYdom mean in group HYsyl 
#        -0.01592911         -0.17443195  

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- HY_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "HYdom") 
                           %>% pull(value), na.rm=T)
    mean_syl_sim <- mean(reshuffled %>% filter(population == "HYsyl") 
                             %>% pull(value), na.rm=T)
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

mean_HYdom <- mean(HY_df %>% filter(population == "HYdom") 
                           %>% pull(value), na.rm=T)
mean_HYsyl <- mean(HY_df %>% filter(population == "HYsyl") 
                             %>% pull(value), na.rm=T)

diff_means_obs <-  mean_HYdom - mean_HYsyl

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#[1] 0



####
# SJ

SJdom <- as_tibble(as.vector(rhod.bs$Fis[,13]))
SJdom <- mutate(SJdom, population = "SJdom")
SJsyl <- as_tibble(as.vector(rhod.bs$Fis[,14]))
SJsyl <- mutate(SJsyl, population = "SJsyl")
SJ_df <- as_tibble(Stack(SJdom, SJsyl))

t.test(value ~ population, data = SJ_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = SJ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#10000 permutations
#t = 0.49935, p-value = 0.6361
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group SJdom mean in group SJsyl 
#         -0.1181607          -0.1281054 


simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- SJ_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "SJdom") 
                           %>% pull(value), na.rm=T)
    mean_syl_sim <- mean(reshuffled %>% filter(population == "SJsyl") 
                             %>% pull(value), na.rm=T)
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)


mean_SJdom <- mean(SJ_df %>% filter(population == "SJdom") 
                           %>% pull(value), na.rm=T)
mean_SJsyl <- mean(SJ_df %>% filter(population == "SJsyl") 
                             %>% pull(value), na.rm=T)

diff_means_obs <-  mean_SJdom  -  mean_SJsyl 

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# 0.6111




########################################################
####
####  Allelic richness 
####
###################################

## BR

BRdom <- as_tibble(as.vector(rhod.ar$Ar[,1]))
BRdom <- mutate(BRdom, population = "BRdom")
BRsyl <- as_tibble(as.vector(rhod.ar$Ar[,2]))
BRsyl <- mutate(BRsyl, population = "BRsyl")
BR_df <- as_tibble(Stack(BRdom, BRsyl))

ggplot(BR_df, aes(x= value, fill=population))+
    ylab("Loci") + xlab("Observed heterozygosities - Ho") +
    geom_histogram(bins = 30, colour = "black") +
    geom_vline(data = filter(BR_df, population == "BRdom"), aes(xintercept = mean(value)),
               size = 1, linetype = "dashed", colour = "red") +
    geom_vline(data = filter(BR_df, population == "BRsyl"), aes(xintercept = mean(value)),
               size = 1, linetype = "dashed", colour = "blue") +
    theme_classic()

t.test(value ~ population, data = BR_df, alternative = "two.sided")
perm.t.test(value ~ population, data = BR_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -8.6904, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group BRdom mean in group BRsyl 
#           1.311909            1.400584 

##############


# Initialize a list to store the simulated test-statistics
simulated_means <- list()

nreps = 10000 # 5000 iterations

for(i in 1:nreps){
    
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- BR_df 
    
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "BRdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "BRsyl") 
                             %>% pull(value))
    
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)


mean_BRdom <- mean(BR_df %>% filter(population == "BRdom") 
                           %>% pull(value), na.rm=T)
mean_BRsyl <- mean(BR_df %>% filter(population == "BRsyl") 
                             %>% pull(value), na.rm=T)


diff_means_obs <- mean_BRdom -  mean_BRsyl


abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# [1]0




###############
#### CE

CEdom <- as_tibble(as.vector(rhod.ar$Ar[,3]))
CEdom <- mutate(CEdom, population = "CEdom")
CEsyl <- as_tibble(as.vector(rhod.ar$Ar[,4]))
CEsyl <- mutate(CEsyl, population = "CEsyl")
CE_df <- as_tibble(Stack(CEdom, CEsyl))

t.test(value ~ population, data = CE_df, alternative = "two.sided")
#t = -4.9597, df = 4794.5, p-value = 7.303e-07
perm.t.test(value ~ population, data = CE_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#	Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -6.4911, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CEdom mean in group CEsyl 
#           1.291144            1.367947

# Initialize a list to store the simulated test-statistics
simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CE_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CEdom") 
                           %>% pull(value), na.rm=T)
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CEsyl") 
                             %>% pull(value), na.rm=T)
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

mean_CEdom <- mean(CE_df %>% filter(population == "CEdom") 
                           %>% pull(value), na.rm=T)
mean_CEsyl <- mean(CE_df %>% filter(population == "CEsyl") 
                             %>% pull(value), na.rm=T)



diff_means_obs <-  1.291144   -   1.367947



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#[1] 0


###############
#### CG

CGdom <- as_tibble(as.vector(rhod.ar$Ar[,5]))
CGdom <- mutate(CGdom, population = "CGdom")
CGsyl <- as_tibble(as.vector(rhod.ar$Ar[,6]))
CGsyl <- mutate(CGsyl, population = "CGsyl")
CG_df <- as_tibble(Stack(CGdom, CGsyl))

t.test(value ~ population, data = CG_df, alternative = "two.sided")


perm.t.test(value ~ population, data = CG_df, alternative = "two.sided", paired = FALSE, nperm = 10000)
#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -3.3981, p-value = 0.0014
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CGdom mean in group CGsyl 
#           1.371226            1.407831 

# Initialize a list to store the simulated test-statistics
simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CG_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CGdom") 
                           %>% pull(value), na.rm=T)
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CGsyl") 
                             %>% pull(value), na.rm=T)
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

mean_CGdom <- mean(CG_df %>% filter(population == "CGdom") 
                           %>% pull(value))
mean_CGsyl <- mean(CG_df %>% filter(population == "CGsyl") 
                             %>% pull(value))



diff_means_obs <- 1.371226  -   1.407831



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
# 0.001


####
# CQ

CQdom <- as_tibble(as.vector(rhod.ar$Ar[,7]))
CQdom <- mutate(CQdom, population = "CQdom")
CQsyl <- as_tibble(as.vector(rhod.ar$Ar[,8]))
CQsyl <- mutate(CQsyl, population = "CQsyl")
CQ_df <- as_tibble(Stack(CQdom, CQsyl))

t.test(value ~ population, data = CQ_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = CQ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = 3.143, p-value = 0.0024
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group CQdom mean in group CQsyl 
#           1.351306            1.314655 


simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- CQ_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "CQdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "CQsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)


mean_CQdom <- mean(CQ_df %>% filter(population == "CQdom") 
                           %>% pull(value))
mean_CQsyl <- mean(CQ_df %>% filter(population == "CQsyl") 
                             %>% pull(value))

diff_means_obs <-  1.351306   -  1.314655 



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#0.001


####
# GL

GLdom <- as_tibble(as.vector(rhod.ar$Ar[,9]))
GLdom <- mutate(GLdom, population = "GLdom")
GLsyl <- as_tibble(as.vector(rhod.ar$Ar[,10]))
GLsyl <- mutate(GLsyl, population = "GLsyl")
GL_df <- as_tibble(Stack(GLdom, GLsyl))

t.test(value ~ population, data = GL_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = GL_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -8.1027, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group GLdom mean in group GLsyl 
#           1.288297            1.381279 

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- GL_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "GLdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "GLsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

mean_GLdom <- mean(GL_df %>% filter(population == "GLdom") 
                           %>% pull(value))
mean_GLsyl <- mean(GL_df %>% filter(population == "GLsyl") 
                             %>% pull(value))

diff_means_obs <- 1.288297  -     1.381279 



abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#0


####
# HY

HYdom <- as_tibble(as.vector(rhod.ar$Ar[,11]))
HYdom <- mutate(HYdom, population = "HYdom")
HYsyl <- as_tibble(as.vector(rhod.ar$Ar[,12]))
HYsyl <- mutate(HYsyl, population = "HYsyl")
HY_df <- as_tibble(Stack(HYdom, HYsyl))

t.test(value ~ population, data = HY_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = HY_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#data:  value by population
#10000 permutations
#t = -6.6018, p-value = 2e-04
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group HYdom mean in group HYsyl 
#           1.211981            1.278213   

simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- HY_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "HYdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "HYsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)

mean_HYdom <- mean(HY_df %>% filter(population == "HYdom") 
                           %>% pull(value))
mean_HYsyl <- mean(HY_df %>% filter(population == "HYsyl") 
                             %>% pull(value))

diff_means_obs <-  1.211981      -  1.278213  

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#[1] 0



####
# SJ

SJdom <- as_tibble(as.vector(rhod.ar$Ar[,13]))
SJdom <- mutate(SJdom, population = "SJdom")
SJsyl <- as_tibble(as.vector(rhod.ar$Ar[,14]))
SJsyl <- mutate(SJsyl, population = "SJsyl")
SJ_df <- as_tibble(Stack(SJdom, SJsyl))

t.test(value ~ population, data = SJ_df, alternative = "two.sided")
# t = 0.68653, df = 4830.8, p-value = 0.4924


perm.t.test(value ~ population, data = SJ_df, alternative = "two.sided", paired = FALSE, nperm = 10000)

#Permutation Two Sample t-test

#10000 permutations
#t = 0.090032, p-value = 0.9217
#alternative hypothesis: true difference in means is not equal to 0
#sample estimates:
#mean in group SJdom mean in group SJsyl 
#           1.216551            1.215721 


simulated_means <- list()
nreps = 10000 # 5000 iterations
for(i in 1:nreps){
    # Create temporary dataframe to permute so we don't modify the original
    reshuffled <- SJ_df 
    # Permute the width column with the 'sample()' function. 
    reshuffled$value <- sample(reshuffled$value, size = nrow(reshuffled), 
                               replace = FALSE)
    # Calculate the means for each sex
    mean_dom_sim <- mean(reshuffled %>% filter(population == "SJdom") 
                           %>% pull(value))
    mean_syl_sim <- mean(reshuffled %>% filter(population == "SJsyl") 
                             %>% pull(value))
    # Calculate to difference between simulated male and female body width 
    # means
    mean_diff_sim <- mean_dom_sim - mean_syl_sim
    # Append simulated mean difference to list
    simulated_means[i] <- mean_diff_sim
}    

# Unlist simulated means list into numeric vector
simulated_means <- unlist(simulated_means)


mean_SJdom <- mean(SJ_df %>% filter(population == "SJdom") 
                           %>% pull(value))
mean_SJsyl <- mean(SJ_df %>% filter(population == "SJsyl") 
                             %>% pull(value))

diff_means_obs <- 1.216551  - 1.215721 

abs_simulated_means <- abs(simulated_means)
abs_diff_means_obs <- abs(diff_means_obs)
exceed_count <- length(abs_simulated_means[abs_simulated_means >= 
                                               abs_diff_means_obs])
p_val <- exceed_count / nreps
#0.9269













```
2 Population with dimicile and sylvatic samples as a group independly of the collection habitat

```{r}

# Remove communities with mixed samples
rhod.mixpop <- popsub(rhod.genind, blacklist=c("SF", "LM", "AH", "ND", "HG", "TC", "VC", "BM", "GA", "EX", "YS", "TR", "SS", "AZ", "NJ", "NT", "RT", "TM"))
# calculate per locus observed and expected heterozygocity {hierfstat}
rhod.mix.bs <- basic.stats(rhod.mixpop, diploid = TRUE, digits = 2)
#calaculate allelic richness
rhod.mix.ar <- allelic.richness(rhod.mixpop)

# Calculate average Observed heterozygosity per population
mean(rhod.mix.bs$Ho[,1])
# Expected heterozygosity Hs
mean(rhod.mix.bs$Hs[,1])
# Calculating mean Fis per population
mean(rhod.mix.bs$Fis[,1], na.rm=T)
# mean allelic richness by pop
mean(rhod.mix.ar$Ar[,1])




```

3. Populations with groups of samples collected at different year and whether there was a insecticide spraying intervention at that year.

```{r}
# 
# Remove communities without samples collected in both habitats (domicile and sylvatic).
rhod.time <- popsub(rhod.genind, blacklist=c("SF", "LM", "AH", "HG", "TC", "VC", "BM", "EX", "YS", "SS", "AZ", "NJ", "NT", "RT", "TM"))

# Change the population map with the approrpiate domicile and sylvatic category.
popmap.time <- as.factor(as.vector(rhod.time@strata$pophabtimetreat))
rhod.time@pop <- popmap.time

# calculate per locus observed and expected heterozygocity {hierfstat}
rhod.time.bs <- basic.stats(rhod.time, diploid = TRUE, digits = 2)
#$overall
#   Ho    Hs    Ht   Dst   Htp  Dstp   Fst  Fstp   Fis  Dest 
# 0.18  0.17  0.22  0.06  0.23  0.06  0.25  0.26 -0.05  0.07 

# calculate allelic richness 
rhod.time.ar <- allelic.richness(rhod.time)

#Global fst

wc(rhod.time)
#$FST
#[1] 0.2771999
#$FIS
#[1] -0.06638569

# Average Observed heterozygosity per population
mean(rhod.time.bs$Ho[,1])
# Expected heterozygosity Hs
mean(rhod.time.bs$Hs[,1])
# Calculating mean Fis per population
mean(rhod.time.bs$Fis[,1], na.rm=T)
# mean allelic reachniss by pop
mean(rhod.time.ar$Ar[,1])

```
4. Calculate alpha diversity for population with only domicile or sylvatic pops

```{r}

# subset samples with only dom or sylvatic pops
rhod.onepop <- popsub(rhod.genind, blacklist=c("BR", "CE", "CG", "CQ","GL", "HY","SJ"))

# calculate basic stats and allelic richness
rhod.onepop.bs <- basic.stats(rhod.onepop, diploid = TRUE, digits = 2)
rhod.onepop.ar <- allelic.richness(rhod.onepop)
# Obs. heterocigocity
mean(rhod.onepop.bs$Ho[,1])
# Exp. Hetero
mean(rhod.onepop.bs$Hs[,1])
#mean Fis
mean(rhod.onepop.bs$Fis[,1], na.rm=T)
# mean allelic richness
mean(rhod.onepop.ar$Ar[,1])

# Calculate loci on HWE
t.one <- seppop(rhod.onepop, pop=rhod.onepop@pop)

length(which(ah[,3]>0.05))/length(ah[,3])

ah <- hw.test(t.one$AH, B=0) # 0.8824451
az <- hw.test(t.one$AZ, B=0) # 0.919279
bm <- hw.test(t.one$BM, B=0)# 1
ex <- hw.test(t.one$EX, B=0) #0.9275078
ga <- hw.test(t.one$GA, B=0) #0.9294671
hg <- hw.test(t.one$HG, B=0) # 0.9275078
lm <- hw.test(t.one$LM, B=0)# 0.8851881
nd <- hw.test(t.one$ND, B=0) # 0.9075235
nj <- hw.test(t.one$NJ, B=0) # 0.9169279
nt <- hw.test(t.one$NT, B=0) # 0.9345611
rt <- hw.test(t.one$RT, B=0)# 0.950627
sf <- hw.test(t.one$SF, B=0)# 0.9416144
ss <- hw.test(t.one$SS, B=0)#  0.8965517
tc <- hw.test(t.one$TC, B=0)# 0.9055643
tm <- hw.test(t.one$TM, B=0)# 0.9467085
tr <- hw.test(t.one$TR, B=0)# 0.9306426
vc <- hw.test(t.one$VC, B=0)# 0.8726489
ys <- hw.test(t.one$YS, B=0)#0.9067398




```


Linkage Disequilibrium using:

https://grunwaldlab.github.io/Population_Genetics_in_R/Linkage_disequilibrium.html

```{r}

# For the whole Loja population




AH.ld <- popsub(rhod.genind, "AH")
ia(AH.ld, sample = 999) 

#         Ia        p.Ia       rbarD        p.rD 
#50.85707633  0.00100000  0.03667747  0.00100000 

AZ.ld <- popsub(rhod.genind, "AZ")
ia(AZ.ld, sample = 999)

#         Ia        p.Ia       rbarD        p.rD 
#8.765237910 0.001000000 0.007699653 0.001000000 

BM.ld <- popsub(rhod.genind, "BM")
ia(BM.ld, sample = 999)
#38.25361272  0.00100000  0.04049636  0.00100000 

BR.ld <- popsub(rhod.genind, "BR")
ia(BR.ld, sample = 999)
#         Ia        p.Ia       rbarD        p.rD 
#21.88864099  0.00100000  0.01374788  0.00100000 

CE.ld <- popsub(rhod.genind, "CE")
ia(CE.ld, sample = 999)

CG.ld <- popsub(rhod.genind, "CG")
ia(CG.ld, sample = 999)


CQ.ld <- popsub(rhod.genind, "CQ")
ia(CQ.ld, sample = 999)

EX.ld <- popsub(rhod.genind, "EX")
ia(EX.ld, sample = 999)


GA.ld <- popsub(rhod.genind, "GA")
ia(GA.ld, sample = 999)

GL.ld <- popsub(rhod.genind, "GL")
ia(GL.ld, sample = 999)

HG.ld <- popsub(rhod.genind, "HG")
ia(HG.ld, sample = 999)


HY.ld <- popsub(rhod.genind, "HY")
ia(HY.ld, sample = 999)


LM.ld <- popsub(rhod.genind, "LM")
ia(LM.ld, sample = 999)



ND.ld <- popsub(rhod.genind, "ND")
ia(ND.ld, sample = 999)


NJ.ld <- popsub(rhod.genind, "NJ")
ia(NJ.ld, sample = 999)


NT.ld <- popsub(rhod.genind, "NT")
ia(NT.ld, sample = 999)


RT.ld <- popsub(rhod.genind, "RT")
ia(RT.ld, sample = 999)



SF.ld <- popsub(rhod.genind, "SF")
ia(SF.ld, sample = 999)


SJ.ld <- popsub(rhod.genind, "SJ")
ia(SJ.ld, sample = 999)


SS.ld <- popsub(rhod.genind, "SS")
ia(SS.ld, sample = 999)


TC.ld <- popsub(rhod.genind, "TC")
ia(TC.ld, sample = 999)


TM.ld <- popsub(rhod.genind, "TM")
ia(TM.ld, sample = 999)



TR.ld <- popsub(rhod.genind, "TR")
ia(TR.ld, sample = 999)



VC.ld <- popsub(rhod.genind, "VC")
ia(VC.ld, sample = 999)



YS.ld <- popsub(rhod.genind, "YS")
ia(YS.ld, sample = 999)





```

Linkage desequilibrium using 

paper: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5972415/
tutorial: https://github.com/AgResearch/GUS-LD
http://htmlpreview.github.io/?https://github.com/AgResearch/GUS-LD/blob/master/doc/Introduction.html

```{r}

install.packages("devtools")
library(devtools)
install_github("tpbilton/GUSbase")
install_github("AgResearch/GUS-LD")



# convert VCF file to an RA file
rafile <- VCFtoRA(infilename = "snps.loja.dplm.recode.vcf", direct = "./")


RAdata <- readRA(rafile = "snps.loja.dplm.recode.vcf.ra.tab", sampthres = 0.01, excsamp = NULL)


urpop <- makeUR(RAobj = RAdata, filter = list(MAF = 0.01, MISS = 0.5, BIN=100, HW = c(-0.05, Inf),
                                              MAXDEPTH = 500), nThreads = 1)

LDres <- GUSLD(URobj = urpop, nClust = 2)

write_tsv(LDres, "LDres.rhodnius.tsv")


## summary stats

# total pairwise LD r2 = 2734291

# Percentage of pairs in LD = 5730
#[1] 0.2095607

# percentage not in LD (i.e., in equilibrium) = 2728561

#[1] 99.79044

### To creat a heat map of pairwise LD r2

LDres.tib <- LDres %>% select(r2, CHROM_SNP1, CHROM_SNP2)

LD.mat = LDres.tib %>% 
  spread(CHROM_SNP1, CHROM_SNP2, fill=0) %>% 
  column_to_rownames(var="CHROM_SNP1") %>% 
  as.matrix


LD.mat <- LDres.tib %>% pivot_wider(LDres.tib, names_from = CHROM_SNP1, values_from = r2)

edge_mat


## LD to whole dataet with poppr


ia(rhod.genind, sample=999)
#Ia       p.Ia      rbarD       p.rD 
#46.0595146  0.0010000  0.0208658  0.0010000 


rhod.pair <- rhod.genind %>% clonecorrect() %>% pair.ia

rhod.pair.df <- as_data_frame(rhod.pair)






write.csv(rhod.pair, "rhod.genind.2552snps.Ia.rbarD.csv")

```

Linkage disequilibrium for pilot 2b rad


```{r}

# set working directory
setwd("C:/Users/quiqu/Dropbox/Recuadoriensis_Analysis/C02_KNNimp/chapt2_analysis")
# linux cluster directory, close when not in cluster
#setwd("~/Dropbox/Recuadoriensis_Analysis/C02_KNNimp")
# Read  in vcf file
pilot.bcgi <- read.genepop("./BcgI.2.90.gen")


# Add population map from sample information file.
bcgi.info <- read.csv("./pilot.strata.bcgi.csv")
popmap.bcgi <- as.factor(as.vector(bcgi.info$struc))
pilot.bcgi@pop <- popmap.bcgi

loja.ld.bcgi <- popsub(pilot.bcgi, "loja")
ia(loja.ld.bcgi, sample = 999) 

#        Ia       p.Ia      rbarD       p.rD 
# 5.70711089 0.00100000 0.03631849 0.00100000 

manabi.ld.bcgi <- popsub(pilot.bcgi, "manabi")
ia(manabi.ld.bcgi, sample = 999) 

#        Ia       p.Ia      rbarD       p.rD 
#4.36314277 0.00100000 0.02090498 0.00100000 


## pairwise

loja.ld.bcgi <- loja.ld.bcgi %>% clonecorrect() %>% pair.ia
loja.ld.bcgi.df <- as_data_frame(loja.ld.bcgi)
write.csv(loja.ld.bcgi.df, "loja.bcgi.df.poppr.csv")

manabi.ld.bcgi.pair <- manabi.ld.bcgi %>% clonecorrect() %>% pair.ia
manabi.ld.bcgi.df <- as_data_frame(manabi.ld.bcgi.pair)
write.csv(manabi.ld.bcgi.df, "manabi.bcgi.df.poppr.csv")



pilot.cspci <- read.genepop("./CspCI.2.90.gen")


# Add population map from sample information file.
cspci.info <- read.csv("./pilot.strata.cspci.csv")
popmap.cspci <- as.factor(as.vector(cspci.info$struc))
pilot.cspci@pop <- popmap.cspci

loja.ld.cspci <- popsub(pilot.cspci, "loja")
ia(loja.ld.cspci, sample = 999) 

#          Ia        p.Ia       rbarD        p.rD 
#40.51305632  0.00100000  0.08518463  0.00100000 

manabi.ld.cspci <- popsub(pilot.cspci, "manabi")
ia(loja.ld.cspci, sample = 999) 

#        Ia       p.Ia      rbarD       p.rD 
#87.8767525  0.0010000  0.1126456  0.0010000 


loja.ld.cspci.pair <- loja.ld.cspci %>% clonecorrect() %>% pair.ia
loja.ld.cspci.df <- as_data_frame(loja.ld.cspci.pair)
write.csv(loja.ld.cspci.df, "loja.cspci.df.poppr.csv")

manabi.ld.cspci.pair <- manabi.ld.cspci %>% clonecorrect() %>% pair.ia
manabi.ld.cspci.df <- as_data_frame(manabi.ld.cspci.pair)
write.csv(manabi.ld.bcgi.df, "manabi.cspcii.df.poppr.csv")

ggplot(loja.ld.bcgi.df, aes(x=rbarD)) +
  geom_histogram() +
  theme_bw() +
  xlab("D")+ 
  ylab("Pairwise SNP comparisons") +
  font("xlab", size= 16, face= "bold") +
  font("ylab", size= 16, face = "bold")

ggplot(loja.ld.cspci.df, aes(x=rbarD)) +
  geom_histogram() +
  theme_bw() +
  xlab("D")+ 
  ylab("Pairwise SNP comparisons") +
  font("xlab", size= 16, face= "bold") +
  font("ylab", size= 16, face = "bold")

ggplot(manabi.ld.bcgi.df, aes(x=rbarD)) +
  geom_histogram() +
  theme_bw() +
  xlab("D")+ 
  ylab("Pairwise SNP comparisons") +
  font("xlab", size= 16, face= "bold") +
  font("ylab", size= 16, face = "bold")

ggplot(manabi.ld.cspci.df, aes(x=rbarD)) +
  geom_histogram() +
  theme_bw() +
  xlab("D")+ 
  ylab("Pairwise SNP comparisons") +
  font("xlab", size= 16, face= "bold") +
  font("ylab", size= 16, face = "bold")





```



AMOVA

1. populations/habitat 272 samples 25 communties
2. populations/habitat 89 samples 7 communities


```{r}

# wc Fst
wc(rhod.genind)

#$FST
#[1] 0.287689
#$FIS
#[1] -0.04400202


# amova with hierstat
rhod.hier <- genind2hierfstat(rhod.genind, pop = rhod.genind@pop)
rhod.varcomp <- varcomp.glob(levels = data.frame(rhod.genind@strata$pop, rhod.genind@strata$hab), rhod.hier[,-1], diploid = TRUE)
rhod.varcomp

# bootstrap for variance components confidence interval and the derived F-statistics
boot.vc(levels = data.frame(rhod.genind@strata$pop, rhod.genind@strata$hab), rhod.hier[,-1],nboot=1000,quant=c(0.025,0.5,0.975))

#signifcance of population
test.between(rhod.hier[,-1], test.lev = rhod.genind@strata$pop, rand.unit = rhod.genind@strata$hab, nperm = 999)
# $p.val
#[1] 0.001001001
# significance of habitat
test.within(rhod.hier[,-1], test = rhod.genind@strata$hab, within = rhod.genind@strata$pop, nperm = 999)
#$p.val
#[1] 0.001001001

# Same analysis for the small dataset with domicile and sylvatic samples.
wc(rhod.domsyl)
#$FST
#[1] 0.2694247
#$FIS
#[1] -0.001796513

rhod.domsyl.hier <- genind2hierfstat(rhod.domsyl, pop = rhod.domsyl@pop)
rhod.domsyl.varcomp <- varcomp.glob(levels = data.frame(rhod.domsyl@strata$pop, rhod.domsyl@strata$hab), rhod.domsyl.hier[,-1], diploid = T)

boot.vc(levels = data.frame(rhod.domsyl@strata$pop, rhod.domsyl@strata$hab), rhod.domsyl.hier[,-1], diploid = T,nboot=1000,quant=c(0.025,0.5,0.975))

test.between(rhod.domsyl.hier[,-1], test.lev = rhod.domsyl@strata$pop, rand.unit = rhod.domsyl@strata$hab, nperm = 999)
test.within(rhod.domsyl.hier[,-1], test = rhod.domsyl@strata$hab, within = rhod.domsyl@strata$pop, nperm = 999)


##### Analysis with time


# amova with hierstat
rhod.hier <- genind2hierfstat(rhod.genind, pop = rhod.genind@pop)
rhod.varcomp <- varcomp.glob(levels = data.frame(rhod.genind@strata$pop, rhod.genind@strata$hab), rhod.genind@strata$coltime, rhod.hier[,-1], diploid = TRUE)
rhod.varcomp

#$overall
#    rhod.genind.strata.pop     rhod.genind.strata.hab rhod.genind.strata.coltime 
#                 151.21427                  -18.69685                   52.19399 
#                       Ind                      Error 
#                 -34.19286                  436.63331 

#$F
#                           rhod.genind.strata.pop rhod.genind.strata.hab rhod.genind.strata.coltime
#Total                                   0.2575386             0.22569531                  0.3145888
#rhod.genind.strata.pop                  0.0000000            -0.04288883                  0.0768393
#rhod.genind.strata.hab                  0.0000000             0.00000000                  0.1148043
#rhod.genind.strata.coltime              0.0000000             0.00000000                  0.0000000
#                                    Ind
#Total                       0.256353705
#rhod.genind.strata.pop     -0.001595917
#rhod.genind.strata.hab      0.039594737
#rhod.genind.strata.coltime -0.084963779


# bootstrap for variance components confidence interval and the derived F-statistics
boot.vc(levels = data.frame(rhod.genind@strata$pop, rhod.genind@strata$hab, rhod.genind@strata$coltime), rhod.hier[,-1],nboot=1000,quant=c(0.025,0.5,0.975))

#signifcance of population
test.between(rhod.hier[,-1], test.lev = rhod.genind@strata$pop, rand.unit = rhod.genind@strata$hab, nperm = 999)
# $p.val
#[1] 0.001001001
# significance of habitat
test.within(rhod.hier[,-1], test = rhod.genind@strata$hab, within = rhod.genind@strata$pop, nperm = 999)
#$p.val
#[1] 0.001001001

test.within(rhod.hier[,-1], test = rhod.genind@strata$coltime, within = rhod.genind@strata$pop, nperm = 999)



#### analysis with time, small dataset

rhod.domsyl.hier <- genind2hierfstat(rhod.domsyl, pop = rhod.domsyl@pop)

rhod.domsyl.varcomp <- varcomp.glob(levels = data.frame(rhod.domsyl@strata$pop, rhod.domsyl@strata$hab, rhod.domsyl@strata$coltime), rhod.domsyl.hier[,-1], diploid = T)

#$overall
#    rhod.domsyl.strata.pop     rhod.domsyl.strata.hab rhod.domsyl.strata.coltime 
#                149.913852                  -2.045492                  30.064515 
#                       Ind                      Error 
#                -18.135604                 436.013845 

#$F
#                           rhod.domsyl.strata.pop rhod.domsyl.strata.hab rhod.domsyl.strata.coltime
#Total                                    0.251613            0.248179927                 0.29863974
#rhod.domsyl.strata.pop                   0.000000           -0.004587362                 0.06283740
#rhod.domsyl.strata.hab                   0.000000            0.000000000                 0.06711687
#rhod.domsyl.strata.coltime               0.000000            0.000000000                 0.00000000
#                                   Ind
#Total                       0.26820122
#rhod.domsyl.strata.pop      0.02216524
#rhod.domsyl.strata.hab      0.02663044
#rhod.domsyl.strata.coltime -0.04339925


boot.vc(levels = data.frame(rhod.domsyl@strata$pop, rhod.domsyl@strata$hab, rhod.domsyl@strata$coltime), rhod.domsyl.hier[,-1], diploid = T,nboot=1000,quant=c(0.025,0.5,0.975))


#$ci
#      H-Total F-rhod.domsyl.strata.pop/Total F-rhod.domsyl.strata.hab/Total
#2.5%   0.2420                         0.2410                         0.2357
#50%    0.2499                         0.2515                         0.2481
#97.5%  0.2573                         0.2627                         0.2605
#      F-rhod.domsyl.strata.coltime/Total F-Ind/Total H-rhod.domsyl.strata.pop
#2.5%                              0.2877      0.2439                   0.1812
#50%                               0.2987      0.2680                   0.1870
#97.5%                             0.3098      0.2912                   0.1925
#      F-rhod.domsyl.strata.hab/rhod.domsyl.strata.pop #F-rhod.domsyl.strata.coltime/rhod.domsyl.strata.pop
#2.5%                                          -0.0135                                              #0.0583
#50%                                           -0.0044                                              #0.0629
#97.5%                                          0.0036                                              #0.0676
#      F-Ind/rhod.domsyl.strata.pop H-rhod.domsyl.strata.hab
#2.5%                       -0.0011                   0.1817
#50%                         0.0219                   0.1878
#97.5%                       0.0463                   0.1938
#      F-rhod.domsyl.strata.coltime/rhod.domsyl.strata.hab F-Ind/rhod.domsyl.strata.hab
#2.5%                                               0.0590                       0.0028
#50%                                                0.0670                       0.0260
#97.5%                                              0.0757                       0.0510
#      H-rhod.domsyl.strata.coltime F-Ind/rhod.domsyl.strata.coltime   Hobs
#2.5%                        0.1699                          -0.0666 0.1748
#50%                         0.1753                          -0.0438 0.1828
#97.5%                       0.1806                          -0.0208 0.1905




test.between(rhod.domsyl.hier[,-1], test.lev = rhod.domsyl@strata$pop, rand.unit = rhod.domsyl@strata$hab, nperm = 999)
test.within(rhod.domsyl.hier[,-1], test = rhod.domsyl@strata$hab, within = rhod.domsyl@strata$pop, nperm = 999)


test.within(rhod.domsyl.hier[,-1], test = rhod.domsyl@strata$coltime, within = rhod.domsyl@strata$pop, nperm = 999)


#$p.val
#[1] 0.001001001

```


 Phylogenic relationship between samples from genetic distances


```{r}

rhod.domsyl.dist <- dist(rhod.domsyl) # euclidean distances from stats package
rhod.domsyl.diss.dist<- diss.dist(rhod.domsyl, percent = FALSE, mat = FALSE) #  distance will be reflected as the number of alleles differing between to individuals from package poppr

rhod.domsyl.dist.gene <- dist.gene(as.data.frame(rhod.domsyl), method = "pairwise", pairwise.deletion = F) # the distance d between two individuals is the number of loci for which they differ, and the associated variance is d(L - d)/L, where L is the number of loci from package ape

rhod.domsyl.d.njtree <- nj(rhod.domsyl.dist)
rhod.domsyl.dd.njtree <- nj(rhod.domsyl.diss.dist)
rhod.domsyl.dg.njtree <- nj(rhod.domsyl.dist.gene)

# Plot tree in R with plot.phylo and change tip or edge/branch colours based on a categories see : http://blog.phytools.org/2017/01/plotting-terminal-edges-of-tree.html

labels <- cbind(indNames(rhod.domsyl), rhod.domsyl@strata$hab)
dom.lab <- subset(labels, subset= labels[,2] == "1")
syl.lab <- subset(labels, subset = labels[,2] == "2")
dom.vec <- as.vector(dom.lab[,1])
syl.vec <- as.vector(syl.lab[,1])
tt<-paintBranches(rhod.domsyl.d.njtree,edge=sapply(dom.vec,match,rhod.domsyl.d.njtree$tip.label),
    state="dom")
tt<-paintBranches(tt, edge=sapply(syl.vec,match,rhod.domsyl.d.njtree$tip.label),
    state="syl")
cols<-setNames(c("blue","red"),c("dom","syl"))
plot.phylo(tt, type="fan", tip.color=cols, edge.width=0.1, cex=0.6, no.margin=F, rotate=60, align.tip.label = TRUE)



plot.phylo(midpoint(rhod.domsyl.d.njtree), type="fan", edge.width=0.1, cex=0.6, no.margin=F, rotate=60, align.tip.label = TRUE)
plot.phylo(midpoint(rhod.domsyl.dd.njtree), type="fan", edge.width=0.1, cex=0.6, no.margin=F, rotate=60, align.tip.label = TRUE)
plot.phylo(midpoint(rhod.domsyl.dg.njtree), type="fan", edge.width=0.1, cex=0.6, no.margin=F, rotate=60, align.tip.label = TRUE)

# Save the different trees created from different genetic distances
write.tree(rhod.domsyl.d.njtree, file="rhod.domsyl.d.tre", append = FALSE, digits = 10, tree.names = FALSE)

write.tree(rhod.domsyl.dd.njtree, file="rhod.domsyl.dd.tre", append = FALSE, digits = 10, tree.names = FALSE)

write.tree(rhod.domsyl.dg.njtree, file="rhod.domsyl.dg.tre", append = FALSE, digits = 10, tree.names = FALSE)


```

Phylogenic relationship between samples and collection sites.


```{r}

rhod.pop.dist <- dist(rhod.genind) # euclidean distances from stats package

# run NJ trees
rhod.pop.d.njtree <- nj(rhod.pop.dist)

# Plot tree in R with plot.phylo and change tip or edge/branch colours based on a categories see : http://blog.phytools.org/2017/01/plotting-terminal-edges-of-tree.html

labels <- cbind(indNames(rhod.genind), rhod.genind@strata$hab)
dom.lab <- subset(labels, subset= labels[,2] == "1")
syl.lab <- subset(labels, subset = labels[,2] == "2")
dom.vec <- as.vector(dom.lab[,1])
syl.vec <- as.vector(syl.lab[,1])

tt<-paintBranches(rhod.pop.d.njtree,edge=sapply(dom.vec,match,rhod.pop.d.njtree$tip.label),
    state="dom")
tt<-paintBranches(tt, edge=sapply(syl.vec,match,rhod.pop.d.njtree$tip.label),
    state="syl")
cols<-setNames(c("blue","red"),c("dom","syl"))
plot.phylo(tt, type="fan", tip.color=cols, edge.width=0.1, cex=0.6, no.margin=F, rotate=60, align.tip.label = TRUE)



plot.phylo(midpoint(rhod.pop.d.njtree), type="fan", edge.width=0.1, cex=0.6, no.margin=F, rotate=60, align.tip.label = TRUE)


# Save the different trees created from different genetic distances
write.tree(rhod.pop.d.njtree, file="./phylo.trees/rhod.pop.d.tre", append = FALSE, digits = 10, tree.names = FALSE)

```


DAPC analysis


```{r}

# Dom vs syl

# Remove communities without samples collected in both habitats (domicile and sylvatic).
rhod.domsyl <- popsub(rhod.genind, blacklist=c("SF", "LM", "AH", "ND", "HG", "TC", "VC", "BM", "GA", "EX", "YS", "TR", "SS", "AZ", "NJ", "NT", "RT", "TM"))
# Change the population map with the approrpiate domicile and sylvatic category.
popmap.hab <- as.factor(as.vector(rhod.domsyl@strata$hab))
rhod.domsyl@pop <- popmap.hab


par(mfrow = c(1,2))
rhod.domsyl.clust <- find.clusters(rhod.domsyl, max.n.clust = 14) # this is the good clustering
100
5

rhod.domsyl.dapc <- dapc(rhod.domsyl, rhod.domsyl.clust$grp) # this is the good dapc
50
3

rhod.domsyl.dapc$grp

cols <- c("royalblue3", "chartreuse3", "black")
myPal <- colorRampPalette(c("blue","gold","red"))

# domsyl

scatter(rhod.domsyl.dapc, grp=rhod.domsyl@strata$hab, posi.da="bottomleft", col=cols, xax=1, yax=3, 
        clab=0.8, solid=0.7, cex=3, pch=19, bg="white", cstar=0, cellipse=0, label = NULL, 
        leg=F, txt.leg=paste("Cluster",1:5))
# time
par(new=TRUE)
df.rhod.domsyl <- data.frame(x = rhod.domsyl.dapc$ind.coord[,1], y = rhod.domsyl.dapc$ind.coord[,3])
noms.rhod.domsyl <- rhod.domsyl@strata$coltime
s.label(dfxy = df.rhod.domsyl, xax=1, yax=2, label=noms.rhod.domsyl,
        clabel=0.7, 
        boxes=FALSE,
        neig = TRUE,
        grid=FALSE, addaxes=TRUE) # do not draw lines or axes in addition to the labels

# Explore population
par(new=TRUE)
df.rhod.domsyl <- data.frame(x = rhod.domsyl.dapc$ind.coord[,1], y = rhod.domsyl.dapc$ind.coord[,3])
noms.rhod.domsyl <- rhod.domsyl@strata$pop
s.label(dfxy = df.rhod.domsyl, xax=1, yax=2, label=noms.rhod.domsyl,
        clabel=0.8, 
        boxes=FALSE,
        neig = TRUE,
        grid=FALSE, addaxes=TRUE)



#####

## Create new data frame from rhod.domsyl.dapc to save DAPC information for using with ggplo2
new.rhod.domsyl.df <- as.data.frame(rhod.domsyl.dapc$ind.coord) # individual coordinates
new.rhod.domsyl.df$Group <- rhod.domsyl.dapc$grp # group assignments
new.rhod.domsyl.df$Pop <- rhod.domsyl@strata$pop # populations
new.rhod.domsyl.df$Hab <- rhod.domsyl@strata$hab # ecotypes

new.r.ds.df <- new.rhod.domsyl.df # create a new data frame to change the legend title and categories 

# Change legend title and factors' names
levels(new.r.ds.df$Hab)[levels(new.r.ds.df$Hab)=="dom"] <- "Domestic"
levels(new.r.ds.df$Hab)[levels(new.r.ds.df$Hab)=="syl"] <- "Sylvatic"
names(new.r.ds.df)[names(new.r.ds.df)=="Hab"]  <- "Ecotype"



# Create a Scatter plot similar to Adegent using PC1 and PC3 coordinates
DAPC13 <-  ggplot(new.r.ds.df, aes(x = LD1, y = LD3)) +
    geom_point(size =22, shape = 20, alpha = 0.6, aes(colour = Ecotype)) +
    geom_text(aes(label = Pop), size = 4) +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    scale_colour_manual(values=cols) + 
    theme_bw() +
  theme(legend.position = "none") +
    xlab("PC 1") +
    ylab("PC 3") +
    font("xlab", size = 25, face = "bold") +
    font("ylab", size = 25, face = "bold") +
   theme(axis.text.x = element_text( size=16,face="bold"), 
         axis.text.y = element_text(size=16,face="bold"))

# Obtain the cumulated variance % for each PCA axis    
temp <- rhod.domsyl.dapc$pca.eig
temp <- 100* cumsum(temp)/sum(temp)

# Create a data frame of this variance and add a PCA.eigenvalue list 1-88
var.df <- as.data.frame(temp) 
var.df$PCA.eig <- as.integer(1:length(temp))
var.df$Retained <- ifelse(var.df$PCA.eig<=50, "Yes", "No") # add a column indicating PCA retained "Yes" or not "No"

# Cumulated variance plot
var.insert <- ggplot(var.df, aes(x=PCA.eig, y=temp))+
  geom_segment(aes(xend = PCA.eig, yend = 0, colour=Retained), size = 1, lineend = "butt") +
  scale_colour_manual(values = c("lightgrey", "black")) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("PCA eigenvalues") +
  ylab("Cumulated variance (%)") +
  font("xlab", size = 16, face = "bold") +
  font("ylab", size = 16, face = "bold") +
   theme(axis.text.x = element_text( size=14,face="bold"), 
         axis.text.y = element_text(size=14,face="bold"))

# Create a data frame of the Discriminant functions
eigenvalues <- rhod.domsyl.dapc$eig
eig.df <- as.data.frame(eigenvalues)
eig.df$DA.eig <- as.factor(1:length(eig.df$eigenvalues))
pc13 <- c("Yes", "No", "Yes", "Null")
pc12 <- c("Yes", "Yes", "No", "Null")
pc23 <- c("No", "Yes", "Yes", "Null")
eig.df$PC13 <- pc13
eig.df$PC12 <- pc12
eig.df$PC23 <- pc23

# Plot discriminant functions eigenvalues
da.insert13<- ggplot(eig.df, aes(x=DA.eig, y=eigenvalues))+
  geom_segment(aes(xend = DA.eig, yend = 0, colour=PC13), 
               size = 14, lineend = "butt") +
  scale_colour_manual(values = c("grey50", "grey86", "black")) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("DA eigenvalues") +
  font("xlab", size = 16, face = "bold") +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
   theme(axis.text.x = element_text( size=14,face="bold"))

# Add inset cumulated varicance plot and Discriminant Analysis eigenvalues
DAPC13 + annotation_custom(ggplotGrob(var.insert), 
                           xmin = -34, 
                           xmax = -18, 
                           ymin = 4.8, 
                           ymax =15) + 
  annotation_custom(ggplotGrob(da.insert13), 
                    xmin = -34, 
                    xmax = -25, 
                    ymin = -5, 
                    ymax = -10)

#####

# Create a Scatter plot similar to Adegent using PC1 and PC2 coordinates
DAPC12 <-  ggplot(new.r.ds.df, aes(x = LD1, y = LD2)) +
    geom_point(size = 20, shape = 20, alpha = 0.7, aes(colour = Ecotype)) +
    geom_text(aes(label = Pop), size = 5) +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    scale_colour_manual(values=cols) + 
    theme_bw() +
  theme(legend.position = "none") +
    xlab("PC 1") +
    ylab("PC 2") +
    font("xlab", size = 20, face = "bold") +
    font("ylab", size = 20, face = "bold")

# Plot discriminant functions eigenvalues
da.insert12<- ggplot(eig.df, aes(x=DA.eig, y=eigenvalues))+
  geom_segment(aes(xend = DA.eig, yend = 0, colour=PC12), 
               size = 14, lineend = "butt") +
  scale_colour_manual(values = c("grey50", "grey86", "black")) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("DA eigenvalues") +
  font("xlab", size = 16, face = "bold") +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

# Add inset cumulated varicance plot and Discriminant Analysis eigenvalues
DAPC12 + annotation_custom(ggplotGrob(var.insert), xmin = -17, xmax = -27, ymin = -20, ymax = -37) +
    annotation_custom(ggplotGrob(da.insert12), xmin = 22, xmax = 32, ymin = -20, ymax = -37)

#####

# Create a Scatter plot similar to Adegent using PC1 and PC2 coordinates
DAPC23 <-  ggplot(new.r.ds.df, aes(x = LD2, y = LD3)) +
    geom_point(size = 20, shape = 20, alpha = 0.7, aes(colour = Ecotype)) +
    geom_text(aes(label = Pop), size = 5) +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    scale_colour_manual(values=cols) + 
    theme_bw() +
  theme(legend.position = "none") +
    xlab("PC 2") +
    ylab("PC 3") +
    font("xlab", size = 20, face = "bold") +
    font("ylab", size = 20, face = "bold")

# Plot discriminant functions eigenvalues
da.insert23<- ggplot(eig.df, aes(x=DA.eig, y=eigenvalues))+
  geom_segment(aes(xend = DA.eig, yend = 0, colour=PC23), 
               size = 14, lineend = "butt") +
  scale_colour_manual(values = c("grey50", "grey86", "black")) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("DA eigenvalues") +
  font("xlab", size = 16, face = "bold") +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

# Add inset cumulated varicance plot and Discriminant Analysis eigenvalues
DAPC23 + annotation_custom(ggplotGrob(var.insert), xmin = -30, xmax = -37, ymin = 8.5, ymax = 14.5) +
    annotation_custom(ggplotGrob(da.insert23), xmin = -30, xmax = -37, ymin = -3, ymax = -9)

#####

# Plot time DAPCs

new.r.ds.df$Coltime <- rhod.domsyl@strata$coltime

# DAPC 13 vs Time
DAPC13.time <-  ggplot(new.r.ds.df, aes(x = LD1, y = LD3)) +
    geom_point(size = 20, shape = 20, alpha = 0.7, aes(colour = Ecotype)) +
    geom_text(aes(label = Coltime), size = 4) +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    scale_colour_manual(values=cols) + 
    theme_bw() +
  theme(legend.position = "none") +
    xlab("PC 1") +
    ylab("PC 3") +
    font("xlab", size = 20, face = "bold") +
    font("ylab", size = 20, face = "bold")

DAPC13.time + annotation_custom(ggplotGrob(var.insert), xmin = 23, xmax = 33, ymin = 6.5, ymax = 16) +
    annotation_custom(ggplotGrob(da.insert13), xmin = 23, xmax = 33, ymin = -3, ymax = -9)


# DAPC 12 vs Time
DAPC12.time <-  ggplot(new.r.ds.df, aes(x = LD1, y = LD2)) +
    geom_point(size = 20, shape = 20, alpha = 0.7, aes(colour = Ecotype)) +
    geom_text(aes(label = Coltime), size = 5) +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    scale_colour_manual(values=cols) + 
    theme_bw() +
  theme(legend.position = "none") +
    xlab("PC 1") +
    ylab("PC 2") +
    font("xlab", size = 20, face = "bold") +
    font("ylab", size = 20, face = "bold")

# Add inset cumulated varicance plot and Discriminant Analysis eigenvalues
DAPC12.time + annotation_custom(ggplotGrob(var.insert), xmin = -17, xmax = -27, ymin = -20, ymax = -37) +
    annotation_custom(ggplotGrob(da.insert12), xmin = 22, xmax = 32, ymin = -20, ymax = -37)

# DAPC 23 vs Time

DAPC23.time <-  ggplot(new.r.ds.df, aes(x = LD2, y = LD3)) +
    geom_point(size = 20, shape = 20, alpha = 0.7, aes(colour = Ecotype)) +
    geom_text(aes(label = Coltime), size = 5) +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    scale_colour_manual(values=cols) + 
    theme_bw() +
  theme(legend.position = "none") +
    xlab("PC 2") +
    ylab("PC 3") +
    font("xlab", size = 20, face = "bold") +
    font("ylab", size = 20, face = "bold")

DAPC23.time + annotation_custom(ggplotGrob(var.insert), xmin = -30, xmax = -37, ymin = 8.5, ymax = 14.5) +
    annotation_custom(ggplotGrob(da.insert23), xmin = -30, xmax = -37, ymin = -3, ymax = -9)



```


Pairwise Fst comparisons among samples from domicile and sylvatic habitat within a communtiy - pairwise.fst adegent using Nei's Fst


```{r}

#("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS")
NBPERM <-999

# BR dom syl pairwise Fst
rhod.BR <- popsub(rhod.genind, blacklist=c( "AH", "AZ", "BM", "CE", "CG", "CQ", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.br <- as.factor(as.vector(rhod.BR@strata$pophab))
rhod.BR@pop <- popmap.br
mat.obs.br <- pairwise.fst(rhod.BR, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.br <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.BR, pop=sample(pop(rhod.BR)), res.type="matrix"), mc.cores = 3)
test.br <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.br[[i]][1,2])), mat.obs.br[1,2], alter="greater")

# CE pairwise comparison
rhod.CE <- popsub(rhod.genind, blacklist=c( "AH", "AZ", "BM", "BR", "CG", "CQ", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.ce <- as.factor(as.vector(rhod.CE@strata$pophab))
rhod.CE@pop <- popmap.ce
mat.obs.ce <- pairwise.fst(rhod.CE, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.ce <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.CE, pop=sample(pop(rhod.CE)), res.type="matrix"), mc.cores = 4)
test.ce <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.ce[[i]][1,2])), mat.obs.ce[1,2], alter="greater")


# CG dom syl pairwise Fst
rhod.CG <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CQ", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.cg <- as.factor(as.vector(rhod.CG@strata$pophab))
rhod.CG@pop <- popmap.cg
mat.obs.cg <- pairwise.fst(rhod.CG, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.cg <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.CG, pop=sample(pop(rhod.CG)), res.type="matrix"), mc.cores = 4)
test.cg <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.cg[[i]][1,2])), mat.obs.cg[1,2], alter="greater")

# CQ pairwise comparisons
rhod.CQ <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CG", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.cq <- as.factor(as.vector(rhod.CQ@strata$pophab))
rhod.CQ@pop <- popmap.cq
mat.obs.cq <- pairwise.fst(rhod.CQ, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.cq <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.CQ, pop=sample(pop(rhod.CQ)), res.type="matrix"), mc.cores = 4)
test.cq <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.cq[[i]][1,2])), mat.obs.cq[1,2], alter="greater")

# GL pairwise comparisons
rhod.GL <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GA", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.gl <- as.factor(as.vector(rhod.GL@strata$pophab))
rhod.GL@pop <- popmap.gl
mat.obs.gl <- pairwise.fst(rhod.GL, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.gl <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.GL, pop=sample(pop(rhod.GL)), res.type="matrix"), mc.cores = 4)
test.gl <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.gl[[i]][1,2])), mat.obs.gl[1,2], alter="greater")

# HY pairwise comparisons
rhod.HY <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GA", "GL", "HG", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.hy <- as.factor(as.vector(rhod.HY@strata$pophab))
rhod.HY@pop <- popmap.hy
mat.obs.hy <- pairwise.fst(rhod.HY, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.hy <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.HY, pop=sample(pop(rhod.HY)), res.type="matrix"), mc.cores = 4)
test.hy <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.hy[[i]][1,2])), mat.obs.hy[1,2], alter="greater")

# SJ pairwise comparisons
rhod.SJ <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GA", "GL", "HG", "HY","LM", "ND", "NJ", "NT", "RT", "SF",  "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.sj <- as.factor(as.vector(rhod.SJ@strata$pophab))
rhod.SJ@pop <- popmap.sj
mat.obs.sj <- pairwise.fst(rhod.SJ, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.sj <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.SJ, pop=sample(pop(rhod.SJ)), res.type="matrix"), mc.cores = 4)
test.sj <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.sj[[i]][1,2])), mat.obs.sj[1,2], alter="greater")

p <- c(0.001, 0.049, 0.182, 0.024, 0.222, 0.005, 0.055)
p.adjust(p, "fdr")
p.adjust(p, "bonferroni")

NBPERM <- 999

# Different year collection group samples
#c("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS")
# Remove communities without samples collected in both habitats (domicile and sylvatic).
rhod.BRT <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "CE", "CG", "CQ", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.brt <- as.factor(as.vector(rhod.BRT@strata$pophabtimetreat))
rhod.BRT@pop <- popmap.brt
mat.obs.brt <- pairwise.fst(rhod.BRT, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.brt <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.BRT, pop=sample(pop(rhod.BRT)), res.type="matrix"), mc.cores = 4)
test.brt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.brt[[i]][1,2])), mat.obs.brt[1,2], alter="greater")

# CG
rhod.CGT <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CQ", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.cgt <- as.factor(as.vector(rhod.CGT@strata$pophabtimetreat))
rhod.CGT@pop <- popmap.cgt
mat.obs.cgt <- pairwise.fst(rhod.CGT, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.cgt <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.CGT, pop=sample(pop(rhod.CGT)), res.type="matrix"), mc.cores = 4)
test.cgt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.cgt[[i]][1,2])), mat.obs.cgt[1,2], alter="greater")

#GL
rhod.GLT <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GA", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.glt <- as.factor(as.vector(rhod.GLT@strata$pophabtimetreat))
rhod.GLT@pop <- popmap.glt
mat.obs.glt <- pairwise.fst(rhod.GLT, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.glt <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.GLT, pop=sample(pop(rhod.GLT)), res.type="matrix"), mc.cores = 4)
test.glt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.glt[[i]][1,2])), mat.obs.glt[1,2], alter="greater")

#SJ
rhod.SJT <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SS", "TC", "TM", "TR", "VC", "YS"))
popmap.sjt <- as.factor(as.vector(rhod.SJT@strata$pophabtimetreat))
rhod.SJT@pop <- popmap.sjt
mat.obs.sjt <- pairwise.fst(rhod.SJT, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.sjt <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.SJT, pop=sample(pop(rhod.SJT)), res.type="matrix"), mc.cores = 4)
test.sjt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.sjt[[i]][1,2])), mat.obs.sjt[1,2], alter="greater")

#ND
rhod.NDT <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GA", "GL", "HG", "HY", "LM",  "NJ", "NT", "RT", "SF", "SJ","SS", "TC", "TM", "TR", "VC", "YS"))
popmap.ndt <- as.factor(as.vector(rhod.NDT@strata$pophabtimetreat))
rhod.NDT@pop <- popmap.ndt
mat.obs.ndt <- pairwise.fst(rhod.NDT, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.ndt <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.NDT, pop=sample(pop(rhod.NDT)), res.type="matrix"), mc.cores = 4)
test.ndt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.ndt[[i]][1,2])), mat.obs.ndt[1,2], alter="greater")

#TR
rhod.TRT <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GA", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ","SS", "TC", "TM", "VC", "YS"))
popmap.trt <- as.factor(as.vector(rhod.TRT@strata$pophabtimetreat))
rhod.TRT@pop <- popmap.trt
mat.obs.trt <- pairwise.fst(rhod.TRT, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.trt <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.TRT, pop=sample(pop(rhod.TRT)), res.type="matrix"), mc.cores = 4)
test.trt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.trt[[i]][1,2])), mat.obs.trt[1,2], alter="greater")

#GA
rhod.GAT <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE", "CG", "CQ", "EX", "GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ","SS", "TC", "TM", "TR", "VC", "YS"))
popmap.gat <- as.factor(as.vector(rhod.GAT@strata$pophabtimetreat))
rhod.GAT@pop <- popmap.gat
mat.obs.gat <- pairwise.fst(rhod.GAT, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.gat <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.GAT, pop=sample(pop(rhod.GAT)), res.type="matrix"), mc.cores = 4)
test.gat <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.gat[[i]][1,2])), mat.obs.gat[1,2], alter="greater")

# Special cases CE, CQ, HY

# CE
rhod.CET <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CG", "CQ", "EX", "GA","GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ","SS", "TC", "TM", "TR", "VC", "YS"))
popmap.cet <- as.factor(as.vector(rhod.CET@strata$pophabtimetreat))
rhod.CET@pop <- popmap.cet
mat.obs.cet <- pairwise.fst(rhod.CET, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.cet <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.CET, pop=sample(pop(rhod.CET)), res.type="matrix"), mc.cores = 4)
test12.cet <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.cet[[i]][1,2])), mat.obs.cet[1,2], alter="greater")
test13.cet <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.cet[[i]][1,3])), mat.obs.cet[1,3], alter="greater")
test23.cet <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.cet[[i]][2,3])), mat.obs.cet[2,3], alter="greater")

# CQ
rhod.CQT <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE","CG", "EX", "GA","GL", "HG", "HY", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ","SS", "TC", "TM", "TR", "VC", "YS"))
popmap.cqt <- as.factor(as.vector(rhod.CQT@strata$pophabtimetreat))
rhod.CQT@pop <- popmap.cqt
mat.obs.cqt <- pairwise.fst(rhod.CQT, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.cqt <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.CQT, pop=sample(pop(rhod.CQT)), res.type="matrix"), mc.cores = 4)
test12.cqt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.cqt[[i]][1,2])), mat.obs.cqt[1,2], alter="greater")
test13.cqt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.cqt[[i]][1,3])), mat.obs.cqt[1,3], alter="greater")
test23.cqt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.cqt[[i]][2,3])), mat.obs.cqt[2,3], alter="greater")


# HY
rhod.HYT <- popsub(rhod.genind, blacklist=c("AH", "AZ", "BM", "BR", "CE","CG", "CQ","EX", "GA","GL", "HG", "LM", "ND", "NJ", "NT", "RT", "SF", "SJ","SS", "TC", "TM", "TR", "VC", "YS"))
popmap.hyt <- as.factor(as.vector(rhod.HYT@strata$pophabtimetreat))
rhod.HYT@pop <- popmap.hyt
mat.obs.hyt <- pairwise.fst(rhod.HYT, res.type="matrix") # this takes about 17 seconds on my computer
mat.perm.hyt <- mclapply(1:NBPERM, function(i) pairwise.fst(rhod.HYT, pop=sample(pop(rhod.HYT)), res.type="matrix"), mc.cores = 4)
test12.hyt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.hyt[[i]][1,2])), mat.obs.hyt[1,2], alter="greater")
test13.hyt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.hyt[[i]][1,3])), mat.obs.hyt[1,3], alter="greater")
test23.hyt <- as.randtest(na.omit(sapply(1:NBPERM, function(i) mat.perm.hyt[[i]][2,3])), mat.obs.hyt[2,3], alter="greater")

p2 <- c(0.001, 0.084, 0.409, 0.022, 0.02, 0.004, 0.022, 0.475, 0.195, 0.06018, 0.017, 0.068, 0.05, 0.006, 0.685, 0.013)

#p.adjust(p2, "bonferroni")
# [1] 0.01600 1.00000 1.00000 0.35200 0.32000 0.06400 0.35200 1.00000 1.00000
#[10] 0.96288 0.27200 1.00000 0.80000 0.09600 1.00000 0.20800

#[1] 0.01600000 0.11200000 0.46742857 0.04400000 0.04400000 0.03200000
# [7] 0.04400000 0.50666667 0.24000000 0.09628800 0.04400000 0.09890909
#[13] 0.08888889 0.03200000 0.68500000 0.04400000


# correlation test between dom and syl geo and gen ditances.

# gen.dist.vec
#[1] 0.05 0.08 0.06 0.06 0.06 0.03 0.10

# geo.vec
#[1] 0.9 1.8 0.1 0.9 3.1 0.5 0.2

#cor.test(gen.dist.vec, geo.vec)
# [1] -0.01014779 p-value 0.98

# 



```

Pairwise Fst comparisons among samples from domicile and sylvatic habitat within a communtiy - using a different approach to Nei's Fst. I will use diveRsity to calculate WC FST sample biased correction.


```{r}


pwGstHendrick.BR.mat <- pairwise_Gst_Hedrick(rhod.BR) #0.12158
pwGstNei.BR.mat <- pairwise_Gst_Nei(rhod.BR) #0.05131588
pwWCFst.BR.mat <- pairwise.fst(rhod.BR) # 0.08480272

pwGstHendrick.CG.mat <- pairwise_Gst_Hedrick(rhod.CG) # 0.04258367
pwGstNei.CG.mat <- pairwise_Gst_Nei(rhod.CG) # 0.01699951
pwWCFst.CG.mat <- pairwise.fst(rhod.CG) # 0.05419255

pwGstHendrick.CQ.mat <- pairwise_Gst_Hedrick(rhod.CQ) # 0.04714119
pwGstNei.CQ.mat <- pairwise_Gst_Nei(rhod.CQ) # 0.01952971
pwWCFst.CQ.mat <- pairwise.fst(rhod.CQ) # 0.05796143

pwGstHendrick.GL.mat <- pairwise_Gst_Hedrick(rhod.GL) # 0.05457195
pwGstNei.GL.mat <- pairwise_Gst_Nei(rhod.GL) # 0.02269748
pwWCFst.GL.mat <- pairwise.fst(rhod.GL) # 0.09762198


pwGstHendrick.CE.mat <- pairwise_Gst_Hedrick(rhod.CE) # 0.03532818
pwGstNei.CE.mat <- pairwise_Gst_Nei(rhod.CE) # 0.01467583
pwWCFst.CE.mat <- pairwise.fst(rhod.CE) # 0.0606314


pwGstHendrick.HY.mat <- pairwise_Gst_Hedrick(rhod.HY) # 0.04974147
pwGstNei.HY.mat <- pairwise_Gst_Nei(rhod.HY) # 0.02203361
pwWCFst.HY.mat <- pairwise.fst(rhod.HY) # 0.06299636


pwGstHendrick.SJ.mat <- pairwise_Gst_Hedrick(rhod.SJ) # -0.003068866
pwGstNei.SJ.mat <- pairwise_Gst_Nei(rhod.SJ) # -0.001349367
pwWCFst.SJ.mat <- pairwise.fst(rhod.SJ) # 0.02568171


```

Isolation by distance analyses separetely for domicile and sylvatic populations


```{r}

# subsample domicile and sylvatic populations

popmap.pophab <- as.factor(as.vector(rhod.genind@strata$pophab))
rhod.pophab<- rhod.genind
rhod.pophab@pop <- popmap.pophab

#c("AHdom", "AZsyl", "BMdom", "BRdom", "BRsyl", "CEdom", "CEsyl", "CGdom", "CGsyl", "CQdom", "CQsyl", "EXdom", "GAdom", "GLdom", "GLsyl", "HGdom", "HYdom", "HYsyl", "LMdom", "NDdom", "NJdom", "NTsyl", "RTsyl", "SFdom", "SJdom", "SJsyl", "SSdom", "TCdom", "TMdom", "TRdom", "VCdom", "YSdom")

rhod.dom <- popsub(rhod.pophab, blacklist=c("AZsyl",  "BRsyl",  "CEsyl", "CGsyl",  "CQsyl",  "GLsyl",  "HYsyl", "NTsyl", "RTsyl", "SJsyl"))

rhod.syl <- popsub(rhod.pophab, blacklist=c("AHdom", "BMdom", "BRdom", "CEdom",  "CGdom",  "CQdom",  "EXdom", "GAdom", "GLdom",  "HGdom", "HYdom",  "LMdom", "NDdom", "NJdom", "SFdom", "SJdom",  "SSdom", "TCdom", "TMdom", "TRdom", "VCdom", "YSdom"))

# Calculate all pairwise Fst and Hendricks Gst matrices.
rhod.dom.pwFst <- pairwise.fst(rhod.dom)
rhod.syl.pwFst <- pairwise.fst(rhod.syl)
rhod.dom.pwGst <- pairwise_Gst_Hedrick(rhod.dom)
rhod.syl.pwGst <- pairwise_Gst_Hedrick(rhod.syl)

write.csv(as.matrix(rhod.dom.pwFst), file = "rhod.dom.pwFst.csv", row.names = T, col.names = T)
write.csv(as.matrix(rhod.syl.pwFst), file = "rhod.syl.pwFst.csv")
write.csv(as.matrix(rhod.dom.pwGst), file = "rhod.dom.pwGst.csv")
write.csv(as.matrix(rhod.syl.pwGst), file = "rhod.syl.pwGst.csv")



# domicile mantel test with Fst matrix
# Pairwise Fst matrix
genmat.dom <- read.csv("rhod.dom.pwFst.csv", header = TRUE, row.names = 1)
Dgen.matrix.dom <- as.matrix(genmat.dom)
Dgen.dom <- as.dist(Dgen.matrix.dom)

#Create a matrix of site locations
sites.dom <- read.csv ("UTMdom.csv")
sites.mat.dom <-as.matrix(sites.dom[c("X", "Y")])
sites.mat.dom
#Get geographic distance in meters
geo.dist.dom <-pointDistance(sites.mat.dom,longlat=F)
Dgeo.dom <-as.dist(geo.dist.dom)

# Mantel test
ibd.dom <- mantel.randtest(Dgen.dom,Dgeo.dom)
#Monte-Carlo test
#Call: mantel.randtest(m1 = Dgen.dom, m2 = Dgeo.dom)
#Observation: 0.4589036 
#Based on 999 replicates
#Simulated p-value: 0.001 
#Alternative hypothesis: greater 
#    Std.Obs Expectation    Variance 
#3.693814571 0.001124974 0.015358924

plot(ibd.dom)
plot(Dgeo.dom, Dgen.dom)
abline(glm(Dgen.dom~Dgeo.dom), col="red",lty=2)


dens.dom <- kde2d(Dgeo.dom,Dgen.dom, n=300)
myPal <- colorRampPalette(c("white","blue","gold", "orange", "red"))
plot(Dgeo.dom, Dgen.dom, pch=20,cex=2)
image(dens, col=transp(myPal(300),.7), add=TRUE)
abline(glm(Dgen.dom~Dgeo.dom), col="red",lty=2)
title("Isolation by distance plot")

# correlograms 
plot(mgram(Dgen.dom, Dgeo.dom))

#$mgram
#           lag ngroup      mantelr  pval         llim          ulim
# [1,]  5358.31     28  0.263278117 0.001  0.203484052  3.583189e-01
# [2,] 16074.93     31  0.251757431 0.007  0.205856603  2.937232e-01
# [3,] 26791.55     48  0.069324515 0.450 -0.001433944  1.409242e-01
# [4,] 37508.17     47 -0.212058853 0.008 -0.318197129 -1.040625e-01
# [5,] 48224.79     39 -0.141297611 0.119 -0.221478670 -4.831148e-02
# [6,] 58941.41     19  0.040210324 0.651 -0.041796127  1.362543e-01
# [7,] 69658.04      4 -0.007217147 0.924 -0.034531010  1.507395e-02
# [8,] 80374.66      7 -0.123823665 0.252 -0.170111033  2.825633e-17
# [9,] 91091.28      7 -0.270417898 0.019 -0.316064780  4.296916e-17


# Sylvatic mantel test with Fst matrix
# Pairwise Fst matrix
genmat.syl <- read.csv("rhod.syl.pwFst.csv", header = TRUE, row.names = 1)
Dgen.matrix.syl <- as.matrix(genmat.syl)
Dgen.syl <- as.dist(Dgen.matrix.syl)

#Create a matrix of site locations
sites.syl <- read.csv ("UTMsyl.csv")
sites.mat.syl <-as.matrix(sites.syl[c("X", "Y")])
sites.mat.syl
#Get geographic distance in meters
geo.dist.syl <-pointDistance(sites.mat.syl,longlat=F)
Dgeo.syl <-as.dist(geo.dist.syl)

# Mantel test
ibd.syl <- mantel.randtest(Dgen.syl,Dgeo.syl)
#Monte-Carlo test
#Call: mantel.randtest(m1 = Dgen.syl, m2 = Dgeo.syl)
#Observation: 0.3116344 
#Based on 999 replicates
#Simulated p-value: 0.043 
#Alternative hypothesis: greater 
#    Std.Obs Expectation    Variance 
#1.838319733 0.000484225 0.028648241 

plot(ibd.syl)
plot(Dgeo.syl, Dgen.syl)
abline(glm(Dgen.syl~Dgeo.syl), col="red",lty=2)

dens.syl <- kde2d(Dgeo.syl,Dgen.syl, n=300)
myPal <- colorRampPalette(c("white","blue","gold", "orange", "red"))
plot(Dgeo.syl, Dgen.syl, pch=20,cex=.5)
image(dens.syl, col=transp(myPal(300),.7), add=TRUE)
abline(glm(Dgen.syl~Dgeo.syl), col="red",lty=2)
title("Isolation by distance plot")

# correlograms 
plot(mgram(Dgen.syl, Dgeo.syl))

#$mgram
 #         lag ngroup      mantelr  pval        llim          ulim
#[1,]  7259.67      6  0.314929838 0.031  0.10654025  5.159934e-01
#[2,] 21779.01      9  0.256346077 0.089  0.14020955  3.940264e-01
#[3,] 36298.35     14 -0.291876131 0.060 -0.47245259 -1.015560e-01
#[4,] 50817.69     11 -0.074106309 0.629 -0.33225450  1.271204e-01
#[5,] 65337.03      1  0.003057791 1.000 -0.03117159  2.767909e-02
#[6,] 79856.37      3 -0.116074236 0.443 -0.26504078  2.460935e-17

```

Mantel test (IBD) + correlograms with the data frame from regression models.

```{r}


gengeo_data <- read_csv("./pair.gen.geo.dist.csv", col_names = T)


##### subset data frame in domestic pairwise populations

dom.df <- filter(gengeo_data, habitat == "Domicile")

dom.mod <- gls(gen_dist ~ geo_dist_km, correlation=corMLPE(form=~pop1.hab+pop2.hab), data=dom.df)

mycol <- c("royalblue3","chartreuse3")

## Add a column with model predicted values and standard errors 
# (to calculate confidence intervals)
dom.df <- bind_cols(dom.df, 
                         setNames(as_tibble(predictSE.gls(dom.mod, dom.df, 
                                                          se.fit = TRUE)[1:2]),
                                  c('fit_link','se_link')))

## Calculate confidence intervals for each correlation from model 
# predicted values and standard errors 
dom.df <- mutate(dom.df,
                         fit_resp  = fit_link,
                         right_upr = fit_link + (2 * se_link),
                         right_lwr = fit_link - (2 * se_link))

ggplot(dom.df, aes(geo_dist_km, gen_dist, shape = habitat, colour = habitat, fill = habitat))+
  geom_point(size=4, alpha=0.7) +
  geom_line(aes(y=fit_resp), size = 1) +
  geom_ribbon(data = dom.df,
              aes(ymin = right_lwr, ymax = right_upr),
              alpha = 0.3) +
  scale_color_manual(values = mycol)+
  scale_fill_manual(values = mycol) +
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Geographic distance (Km)") +
  ylab(bquote('Genetic distance ('*F[ST]*')')) +
  font("xlab", size = 18, face = "bold") +
  font("ylab", size = 18, face = "bold")


ibd.dom <- ggplot(dom.df, aes(geo_dist_km, gen_dist, shape = habitat, colour = habitat, fill = habitat))+
  geom_point(size=4, alpha=0.7) +
  geom_smooth(method="glm", alpha=0.5, size=2) +
  scale_color_manual(values = mycol)+
  scale_fill_manual(values = mycol) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Geographic distance (Km)") +
  ylab(bquote('Genetic distance ('*F[ST]*')')) +
  font("xlab", size = 14, face = "bold") +
  font("ylab", size = 18, face = "bold")

# Create a mantel correlogram inset

mgram.dom <- mgram(Dgen.dom, Dgeo.dom)

mgram.dom.df <- as_data_frame(mgram.dom$mgram)
mgram.dom.df <- mutate(mgram.dom.df, lag.km = lag/1000)
cols = c("black", "white")

mgram.inset.dom <- ggplot(mgram.dom.df, aes(lag.km, mantelr)) +
  geom_line(colour="royalblue3", size=1.5) +
  geom_point(aes(fill = cut(pval, c(0, 0.05, 1))), size = 5, pch=21) +
  scale_fill_manual(values = cols) +
  theme_bw() +
  theme(legend.position = "none")  +
  xlab("Geographic distance (Km)") +
  ylab(bquote('Mantel correlation ('*r[m]*')')) +
  ylim(c(-0.28,0.32)) +
   theme(axis.text.x = element_text( size=18,face="bold"), 
         axis.text.y = element_text(size=18,face="bold")) + 
   theme(axis.title.x = element_text(size=25, face="bold"),
         axis.title.y = element_text( size=28, face="bold")) +
  geom_hline(yintercept=0, lty="dashed")


ibd.dom + annotation_custom(ggplotGrob(mgram.inset.dom), xmin = 65, xmax = 100, ymin = 0.01, ymax = 0.13)

##### Sylvatic data frame, mantel test (IBD) and correlogram

syl.df <- filter(gengeo_data, habitat == "Sylvatic")

ibd.syl <- ggplot(syl.df, aes(geo_dist_km, gen_dist, shape = habitat, colour = habitat, fill = habitat))+
  geom_point(size=4, alpha=0.7) +
  geom_smooth(method="glm", alpha=0.5, size=2) +
  scale_color_manual(values = "chartreuse3")+
  scale_fill_manual(values = "chartreuse3") +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Geographic distance (Km)") +
  ylab(bquote('Genetic distance ('*F[ST]*')')) +
  font("xlab", size = 14, face = "bold") +
  font("ylab", size = 18, face = "bold")

# Create a mantel correlogram inset

mgram.syl <- mgram(Dgen.syl, Dgeo.syl)

mgram.syl.df <- as_data_frame(mgram.syl$mgram)
mgram.syl.df <- mutate(mgram.syl.df, lag.km = lag/1000)
cols = c("black", "white")

mgram.inset.syl <- ggplot(mgram.syl.df, aes(lag.km, mantelr)) +
  geom_line(colour="chartreuse3", size=1.5) +
  geom_point(aes(fill = cut(pval, c(0, 0.05, 1))), size = 5, pch=21) +
  scale_fill_manual(values = cols) +
  theme_bw() +
  theme(legend.position = "none")  +
  xlab("Geographic distance (Km)") +
  ylab(bquote('Mantel correlation ('*r[m]*')')) +
  ylim(c(-0.30,0.32)) +
   theme(axis.text.x = element_text( size=18,face="bold"), 
         axis.text.y = element_text(size=18,face="bold")) + 
   theme(axis.title.x = element_text(size=25, face="bold"),
         axis.title.y = element_text( size=28, face="bold")) +
  geom_hline(yintercept=0, lty="dashed") +
  theme(axis.title.y = element_blank())


ibd.syl + annotation_custom(ggplotGrob(mgram.inset.syl), xmin = 55, xmax = 92, ymin = 0.04, ymax = 0.13)


### plot mantel correlograms dom and syl side by side

plot_grid(mgram.inset.dom,  mgram.inset.syl, align = "h", nrow  = 2)



### Populations IBD

genmat.pop <- read.csv("./rhod.pop.pwFst.csv", header = TRUE, row.names = 1)
Dgen.matrix.pop <- as.matrix(genmat.pop)
Dgen.pop <- as.dist(Dgen.matrix.pop)
sites.pop <- read.csv("./UTMpops.csv")
sites.mat.pop <- as.matrix(sites.pop[c("X", "Y")])
geo.dist.pop <- pointDistance(sites.mat.pop, longlat=F)
D.geo.pop <- as.dist(geo.dist.pop)
D.geo.pop

# Create a tibble data frames of the pairwise genetic and geographic distances between communties.
df.Dgen.pop <- as_tibble(df.Dgen.pop <- melt(as.matrix(Dgen.pop), varnames = c("row", "col")))
df.Dgeo.pop <- as_tibble(df.Dgeo.pop <- melt(as.matrix(D.geo.pop), varnames = c("row", "col")))
# Add the value of pairwise geographic distances to the pairwise geentic distances dataframe
df.Dgengeo.pop <- df.Dgen.pop %>% right_join(df.Dgeo.pop, by=c("row","col"))

# Remove zero values for appropriate regression
new.df.Dgengeo.pop <- df.Dgengeo.pop %>% filter(value.x>0, value.y>0) # remove zero values

#rename coloumns names
new.df.Dgengeo.pop <- new.df.Dgengeo.pop %>% rename(pop1 = row, pop2 = col, gen_dist = value.x, geo_dist = value.y)

write_csv(new.df.Dgengeo.pop, "pwFst.pop.df.csv")

new.df.Dgengeo.pop <- mutate(new.df.Dgengeo.pop, geo.dist.km = geo_dist/1000)


ibd.pop <- ggplot(new.df.Dgengeo.pop, aes(geo.dist.km, gen_dist))+
  geom_point(size=4, alpha=0.7) +
  geom_smooth(method = "glm", alpha=0.5, size=2) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Geographic distance (Km)") +
  ylab(bquote('Genetic distance ('*F[ST]*')')) +
  font("xlab", size = 14, face = "bold") +
  font("ylab", size = 18, face = "bold")


# Create a mantel correlogram inset

mgram.pop <- mgram(Dgen.pop, D.geo.pop)

mgram.pop.df <- as_data_frame(mgram.pop$mgram)
mgram.pop.df <- mutate(mgram.pop.df, lag.km = lag/1000)
cols = c("black", "white")

mgram.inset.pop <- ggplot(mgram.pop.df, aes(lag.km, mantelr)) +
  geom_line() +
  geom_point(aes(fill = cut(pval, c(0, 0.05, 1))), size = 4, pch=21) +
  scale_fill_manual(values = cols) +
  theme_bw() +
  theme(legend.position = "none")  +
  xlab("Geographic distance (Km)") +
  ylab(bquote('Mantel correlation ('*r[m]*')')) +
  font("xlab", size = 14, face = "bold") +
  font("ylab", size = 18, face = "bold") +
  geom_hline(yintercept=0, lty="dashed")


ibd.pop + annotation_custom(ggplotGrob(mgram.inset.pop), xmin = 70, xmax = 100, ymin = 0, ymax = 0.12)


```

Prepare dataframe for regression models 

```{r}

plot(Dgeo.dom, Dgen.dom, col="royalblue3", cex=1, pch=19)
points(Dgeo.syl, Dgen.syl, col="chartreuse3", cex=1, pch=19)
abline(lm(Dgen.dom~Dgeo.dom), col="royalblue3", lty=2)
abline(lm(Dgen.syl~Dgeo.syl), col="chartreuse3", lty=2)

# Create a tibble data frames of the pairwise genetic and geographic distances between communties.
df.Dgen.dom <- as_tibble(df.Dgen.dom <- melt(as.matrix(Dgen.dom), varnames = c("row", "col")))
df.Dgeo.dom <- as_tibble(df.Dgeo.dom <- melt(as.matrix(Dgeo.dom), varnames = c("row", "col")))
# Add the value of pairwise geographic distances to the pairwise geentic distances dataframe
df.Dgengeo <- df.Dgen.dom %>% right_join(df.Dgeo.dom, by=c("row","col"))
# Add a new column with a cotegorical variable called Habitat
df.Dgengeo <- mutate(df.Dgengeo, habitat = "Domicile")

df.Dgen.syl <- as_tibble(df.Dgen.syl <- melt(as.matrix(Dgen.syl), varnames = c("row", "col")))
df.Dgeo.syl <- as_tibble(df.Dgeo.syl <- melt(as.matrix(Dgeo.syl), varnames = c("row", "col")))
df.Dgengeo.syl <- df.Dgen.syl %>% right_join(df.Dgeo.syl, by=c("row","col"))
df.Dgengeo.syl <- mutate(df.Dgengeo.syl, habitat = "Sylvatic") 

# Stack domicile and sylvatic data frames on top of each other.
df.Dgengeo.domsyl <- as_tibble(Stack(df.Dgengeo, df.Dgengeo.syl))
# Remove zero values for appropriate regression
new.df.Dgengeo.domsyl <- df.Dgengeo.domsyl %>% filter(value.x>0, value.y>0) # remove zero values

new.df.Dgengeo.domsyl <- new.df.Dgengeo.domsyl %>% rename(pop1 = row, pop2 = col, gen_dist = value.x, geo_dist = value.y)

write_csv(new.df.Dgengeo.domsyl, "pair.gen.geo.dist.csv")

```


Linear regression model - ancova


```{r}

mycol <- c("royalblue3","chartreuse3")

wcFst <-ggplot(new.df.Dgengeo.domsyl, aes(value.y, value.x, shape=habitat, colour= habitat, fill=habitat))+
  geom_point(size=4, alpha=0.7) +
  geom_smooth(method="lm", alpha=0.5, size=2) + 
  scale_color_manual(values =mycol)+
  scale_fill_manual(values = mycol) +
  theme_bw() + 
  xlab("Geographic distance (m)") +
  ylab(bquote('Genetic distance ('*F[ST]*')'))

rousset.Fst <-ggplot(new.df.Dgengeo.domsyl, aes(value.y, (value.x/(1-value.x)), shape=habitat, colour= habitat, fill=habitat))+
  geom_point(size=4, alpha=0.7) +
  geom_smooth(method="lm", alpha=0.5, size=2) + 
  scale_color_manual(values =mycol)+
  scale_fill_manual(values = mycol) +
  theme_bw() + 
  xlab("Geographic distance (m)") +
  ylab(bquote('Genetic distance ('*F[ST]/(1-F[ST])*')'))

domsyl.m1 <- lm(formula=value.x~value.y*habitat, data = new.df.Dgengeo.domsyl)
domsyl.m2 <- lm(formula=value.x~value.y+habitat, data = new.df.Dgengeo.domsyl)
domsyl.m3 <- lm(formula=value.x~value.y, data = new.df.Dgengeo.domsyl)

summary(domsyl.m1)
summary(domsyl.m2)
summary(domsyl.m3)

anova(domsyl.m2)
#Analysis of Variance Table

#Model 1: value.x ~ value.y + habitat
#Model 2: value.x ~ value.y * habitat
#  Res.Df    RSS Df Sum of Sq      F Pr(>F)
#1    549 2.5074                           
#2    548 2.4999  1 0.0075419 1.6533 0.1991

domsyl.r.m1 <- lm(formula=(value.x/(1-value.x))~value.y*habitat, data = new.df.Dgengeo.domsyl)
domsyl.r.m2 <- lm(formula=(value.x/(1-value.x))~value.y+habitat, data = new.df.Dgengeo.domsyl)
summary(domsyl.r.m1)
summary(domsyl.r.m2)
anova(domsyl.r.m2, domsyl.r.m1 )

#Analysis of Variance Table

#Model 1: (value.x/(1 - value.x)) ~ value.y + habitat
#Model 2: (value.x/(1 - value.x)) ~ value.y * habitat
#    Res.Df    RSS Df Sum of Sq      F Pr(>F)
#1    549 6.3538                           
#2    548 6.3275  1  0.026342 2.2813 0.1315

```

Linear mixed-effects model 


```{r}

gengeo_data <- read_csv("./pair.gen.geo.dist.csv", col_names = T)

gengeo_data$pop1.hab <- paste(gengeo_data$habitat, gengeo_data$pop1, sep = "-")
gengeo_data$pop2.hab <- paste(gengeo_data$habitat, gengeo_data$pop2, sep = "-")
gengeo_data$geo_dist_km <- as.double(round(gengeo_data$geo_dist/1000, digits =1))

lmemod1 <- lme(gen_dist ~ geo_dist, random = ~geo_dist|habitat, correlation=corMLPE(form=~pop1+pop2|habitat), data=gengeo_data)

lmemod2 <- gls(gen_dist ~ geo_dist_km, correlation=corMLPE(form=~pop1.hab+pop2.hab), data=gengeo_data)

lmemod3 <- gls(gen_dist ~ geo_dist_km + habitat, correlation=corMLPE(form=~pop1.hab+pop2.hab), data=gengeo_data)
#summary(lmemod3)
#Generalized least squares fit by REML
#  Model: gen_dist ~ geo_dist_km + habitat 
#  Data: gengeo_data 
 #       AIC       BIC   logLik
#  -1851.935 -1830.394 930.9674

#Correlation Structure: corMLPE
# Formula: ~pop1.hab + pop2.hab 
# Parameter estimate(s):
#      rho 
#0.3456614 

#Coefficients:
#                     Value  Std.Error   t-value p-value
#(Intercept)      0.1084149 0.01834888  5.908532  0.0000
#geo_dist_km      0.0019010 0.00011285 16.845851  0.0000
#habitatSylvatic -0.0042290 0.03218032 -0.131416  0.8955

# Correlation: 
#                (Intr) g_dst_
#geo_dist_km     -0.219       
#habitatSylvatic -0.542 -0.005

#Standardized residuals:
 #      Min         Q1        Med         Q3        Max 
#-2.1432625 -0.6319959 -0.2008772  0.6184771  2.2171929 

#Residual standard error: 0.07103711 
#Degrees of freedom: 552 total; 549 residual
rsquared(lmemod3)
# 0.2564392

anova(lmemod3)
#Denom. DF: 549 
#            numDF   F-value p-value
#(Intercept)     1 139.49374  <.0001
#geo_dist_km     1 283.76738  <.0001
#habitat         1   0.01727  0.8955

lmemod4 <- gls(gen_dist ~ geo_dist_km * habitat, correlation=corMLPE(form=~pop1.hab+pop2.hab), data=gengeo_data)
#Generalized least squares fit by REML
#  Model: gen_dist ~ geo_dist_km * habitat 
#  Data: gengeo_data 
#        AIC      BIC   logLik
#  -1866.328 -1840.49 939.1638

#Correlation Structure: corMLPE
# Formula: ~pop1.hab + pop2.hab 
# Parameter estimate(s):
#      rho 
#0.3771037 

#Coefficients:
#                                  Value  Std.Error   t-value p-value
#(Intercept)                  0.11696273 0.02062838  5.669991  0.0000
#geo_dist_km                  0.00166108 0.00011759 14.125935  0.0000
#habitatSylvatic             -0.07167028 0.03812622 -1.879816  0.0607
#geo_dist_km:habitatSylvatic  0.00182716 0.00031929  5.722646  0.0000

# Correlation: 
#                            (Intr) g_dst_ hbttSy
#geo_dist_km                 -0.203              
#habitatSylvatic             -0.541  0.110       
#geo_dist_km:habitatSylvatic  0.075 -0.368 -0.309

#Standardized residuals:
#       Min         Q1        Med         Q3        Max 
#-2.7889540 -0.6507613 -0.1701553  0.6010450  2.0827449 

#Residual standard error: 0.07684065 
#Degrees of freedom: 552 total; 548 residual

intervals(lmemod4)

#Approximate 95% confidence intervals

# Coefficients:
#                                   lower         est.       upper
#(Intercept)                  0.076442354  0.116962734 0.157483115
#geo_dist_km                  0.001430100  0.001661085 0.001892069
#habitatSylvatic             -0.146561704 -0.071670282 0.003221139
#geo_dist_km:habitatSylvatic  0.001199983  0.001827155 0.002454328
#attr(,"label")
#[1] "Coefficients:"

# Correlation structure:
#        lower      est.     upper
#rho 0.3200792 0.3771037 0.4205414
#attr(,"label")
#[1] "Correlation structure:"

# Residual standard error:
#     lower       est.      upper 
#0.06291271 0.07684065 0.09385203 



write.csv(gengeo_data, file = "gengeo_data.csv")

mycol <- c("royalblue3","chartreuse3")


## add fit and se.fit on the **link** scale
gengeo_data <- bind_cols(gengeo_data, setNames(as_tibble(predictSE.gls(lmemod5, gengeo_data, se.fit = TRUE)[1:2]),
                                                     c('fit_mod5','se_mod5')))



## create the interval and backtransform
gengeo_data <- mutate(gengeo_data,
                         fit_resp  = fit_mod5,
                         right_upr = fit_mod5 + (2 * se_mod5),
                         right_lwr = fit_mod5 - (2 * se_mod5))


# with ggeffects

plot(ggemmeans(lmemod4, terms = c("geo_dist_km", "habitat")))



ggplot(gengeo_data, aes(geo_dist_km, gen_dist, shape=habitat, colour= habitat, fill=habitat))+
  geom_point(size=4) +
  scale_shape_manual(values=c(21,23)) +
  geom_line(aes(y=fit_resp), size = 1) +
  geom_ribbon(data = gengeo_data,
              aes(ymin = right_lwr, ymax = right_upr),
              alpha = 0.2,  colour = NA) +
  scale_color_manual(values = mycol) +
  scale_fill_manual( values = mycol) +
  theme_bw() + 
  xlab("Geographic distance (Km)") +
  ylab(bquote('Genetic distance ('*F[ST]*')')) +
  guides(fill = guide_legend(ncol = 1))  +
   theme(axis.text.x = element_text( size=18,face="bold"), 
         axis.text.y = element_text(size=18,face="bold")) + 
   theme(axis.title.x = element_text(size=25, face="bold"),
         axis.title.y = element_text( size=28, face="bold")) +
  theme(legend.position = "none") 




# Plot residuals
residplot(lmemod4)


```

Domicile mantel test with Hendrick Gst matrix
```{r}

genmat.dom <- read.csv("./rhod.dom.pwGst.csv", header = TRUE, row.names = 1)
Dgen.matrix.dom <- as.matrix(genmat.dom)
Dgen.dom <- as.dist(Dgen.matrix.dom)

#Create a matrix of site locations
sites.dom <- read.csv ("UTMdom.csv")
sites.mat.dom <-as.matrix(sites.dom[c("X", "Y")])
sites.mat.dom
#Get geographic distance in meters
geo.dist.dom <-pointDistance(sites.mat.dom,longlat=F)
Dgeo.dom <-as.dist(geo.dist.dom)

# Mantel test
ibd.dom <- mantel.randtest(Dgen.dom,Dgeo.dom)

#Monte-Carlo test
#Call: mantel.randtest(m1 = Dgen.dom, m2 = Dgeo.dom)

#Observation: 0.4717461 

#Based on 999 replicates
#Simulated p-value: 0.001 
#Alternative hypothesis: greater 

#     Std.Obs  Expectation     Variance 
# 3.819369744 -0.005170374  0.015591979

plot(ibd.dom)
plot(Dgeo.dom, Dgen.dom)
abline(glm(Dgen.dom~Dgeo.dom), col="red",lty=2)

dens.dom <- kde2d(Dgeo.dom,Dgen.dom, n=300)
myPal <- colorRampPalette(c("white","blue","gold", "orange", "red"))
plot(Dgeo.dom, Dgen.dom, pch=20,cex=.5)
image(dens.dom, col=transp(myPal(300),.7), add=TRUE)
abline(glm(Dgen.dom~Dgeo.dom), col="red",lty=2)


# correlograms 
plot(mgram(Dgen.dom, Dgeo.dom))

#$mgram
#           lag ngroup     mantelr  pval        llim          ulim
# [1,]  5358.31     28  0.27667103 0.001  0.21724984  3.901866e-01
# [2,] 16074.93     31  0.24023134 0.005  0.18713014  2.939219e-01
# [3,] 26791.55     48  0.12419721 0.182  0.03124524  2.220211e-01
# [4,] 37508.17     47 -0.23232491 0.002 -0.33317557 -1.120278e-01
# [5,] 48224.79     39 -0.19201352 0.022 -0.26257176 -1.083473e-01
# [6,] 58941.41     19  0.04566118 0.585 -0.03674024  1.259654e-01
# [7,] 69658.04      4 -0.01115473 0.887 -0.03884384  1.110006e-02
# [8,] 80374.66      7 -0.11905397 0.273 -0.15630914  3.032338e-17
# [9,] 91091.28      7 -0.25549826 0.021 -0.30267904  5.719991e-18




# Sylvatic mantel test with Hendricks Gst matrix.

# Pairwise Fst matrix
genmat.syl <- read.csv("rhod.syl.pwGst.csv", header = TRUE, row.names = 1)
Dgen.matrix.syl <- as.matrix(genmat.syl)
Dgen.syl <- as.dist(Dgen.matrix.syl)

#Create a matrix of site locations
sites.syl <- read.csv ("UTMsyl.csv")
sites.mat.syl <-as.matrix(sites.syl[c("X", "Y")])
sites.mat.syl
#Get geographic distance in meters
geo.dist.syl <-pointDistance(sites.mat.syl,longlat=F)
Dgeo.syl <-as.dist(geo.dist.syl)

# Mantel test
ibd.syl <- mantel.randtest(Dgen.syl,Dgeo.syl)

plot(ibd.syl)
plot(Dgeo.syl, Dgen.syl)
abline(glm(Dgen.syl~Dgeo.syl), col="red",lty=2)

dens.syl <- kde2d(Dgeo.syl,Dgen.syl, n=300)
myPal <- colorRampPalette(c("white","blue","gold", "orange", "red"))
plot(Dgeo.syl, Dgen.syl, pch=20,cex=.5)
image(dens.syl, col=transp(myPal(300),.7), add=TRUE)
abline(glm(Dgen.syl~Dgeo.syl), col="red",lty=2)
title("Isolation by distance plot")

# correlograms 
plot(mgram(Dgen.syl, Dgeo.syl))

#$mgram
#          lag ngroup     mantelr  pval        llim          ulim
#[1,]  7259.67      6  0.37529144 0.009  0.10821254  5.702839e-01
#[2,] 21779.01      9  0.16849124 0.338  0.02946252  3.252022e-01
#[3,] 36298.35     14 -0.24897970 0.099 -0.43987982 -4.853562e-02
#[4,] 50817.69     11 -0.10678453 0.485 -0.34985600  1.068986e-01
#[5,] 65337.03      1 -0.03667396 0.965 -0.07074864  8.330271e-17
#[6,] 79856.37      3 -0.07906545 0.657 -0.24578835  1.194972e-02




```

Correlation Fst dom syl pairs vs geographic distances


```{r}

gen.dist <- c(0.05, 0.08, 0.06, 0.06, 0.06, 0.03, 0.10)
geo.dist <- c(0.892065, 1.853765, 0.1040141, 0.9685069, 3.188121, 0.582964, 0.2456451)
pop <- c("CG", "BR", "CE", "CQ", "HY", "SJ", "GL")
domsyl.df <- data.frame(gen.dist, geo.dist, pop)

ggplot(domsyl.df, aes(x = geo.dist, y= gen.dist)) + 
  geom_point() + 
  geom_smooth(method = "lm", alpha=0.5, size=2) + 
  theme_bw() + 
  xlab("Geographic distance (Km)") + 
  ylab(bquote('Genetic distance ('*F[ST]*')'))

ibd.domsyl <- mantel(formula = as.matrix(gen.dist) ~ as.matrix(geo.dist), data = domsyl.df)

perm.cor.test(domsyl.df$gen.dist, domsyl.df$geo.dist, alternative = "greater", num.sim = 10000)


```

DAPC with RF SNPS

```{r}

rhod.rf.vcf <- read.vcfR("./snps.loja.KNNimpR3.rf.recode.vcf")

# Convert vcf to genind object
rhod.rf.genind <- vcfR2genind(rhod.rf.vcf, sep = "/")

# Add population map from sample information file.
rhod.info <- read.csv("./C02KNNimpR3_strata.csv")
popmap <- as.factor(as.vector(rhod.info$pop))
rhod.rf.genind@pop <- popmap
rhod.rf.genind

strata(rhod.rf.genind) <- rhod.info
rhod.rf.genind

# Remove communities without samples collected in both habitats (domicile and sylvatic).
rhod.rf.domsyl <- popsub(rhod.rf.genind, blacklist=c("SF", "LM", "AH", "ND", "HG", "TC", "VC", "BM", "GA", "EX", "YS", "TR", "SS", "AZ", "NJ", "NT", "RT", "TM"))
# Change the population map with the approrpiate domicile and sylvatic category.
popmap.hab <- as.factor(as.vector(rhod.domsyl@strata$hab))
rhod.rf.domsyl@pop <- popmap.hab



par(mfrow = c(1,2))
rhod.rf.domsyl.clust <- find.clusters(rhod.rf.domsyl, max.n.clust = 20) # this is the good clustering
15
10

rhod.domsyl.dapc <- dapc(rhod.rf.domsyl, rhod.rf.domsyl.clust$grp) # this is the good dapc
10
3
rhod.rf.domsyl.dapc$grp

cols <- c("royalblue3", "chartreuse3")
myPal <- colorRampPalette(c("blue","gold","red"))

# domsyl

scatter(rhod.domsyl.dapc, grp=rhod.rf.domsyl@strata$hab, posi.da="bottomleft", col=cols, xax=1, yax=2, 
        clab=0.8, solid=0.7, cex=3, pch=19, bg="white", cstar=0, cellipse=0, label = NULL, 
        leg=F, txt.leg=paste("Cluster",1:5))
# time
par(new=TRUE)
df.rhod.domsyl <- data.frame(x = rhod.domsyl.dapc$ind.coord[,1], y = rhod.domsyl.dapc$ind.coord[,3])
noms.rhod.domsyl <- rhod.domsyl@strata$coltime
s.label(dfxy = df.rhod.domsyl, xax=1, yax=2, label=noms.rhod.domsyl,
        clabel=0.7, 
        boxes=FALSE,
        neig = TRUE,
        grid=FALSE, addaxes=TRUE) # do not draw lines or axes in addition to the labels

# Explore population
par(new=TRUE)
df.rhod.domsyl <- data.frame(x = rhod.domsyl.dapc$ind.coord[,1], y = rhod.domsyl.dapc$ind.coord[,3])
noms.rhod.domsyl <- rhod.domsyl@strata$pop
s.label(dfxy = df.rhod.domsyl, xax=1, yax=2, label=noms.rhod.domsyl,
        clabel=0.8, 
        boxes=FALSE,
        neig = TRUE,
        grid=FALSE, addaxes=TRUE)







```
