---
title: "Rhodnius resistanceGA optimisation"
author: "Enrique Hernandez"
date: "15/12/2020"
output:
  pdf_document: default
  html_document: default
---


R packages:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(rgdal)
library(gdistance)
library(ResistanceGA)

# Read working directories for each cluster/PC

# Lenovo windows
setwd("C:/Users/quiqu/Dropbox (Personal)/Rhodnius_project_2019/LandGen_analysis/R_scripts/ResistanceGA_Loja/all_comb")

write.dir <-"./Rga_gdistance_151220/"


# getafix directory
setwd("~/Dropbox/Rhodnius_project_2019/LandGen_analysis/R_scripts/ResistanceGA_Loja/all_comb")
# getafix write results
write.dir <-"./"


# Orion directory
setwd("/shared2/salmosim/luis2/ecuadoriensis/landgen_analysis/ResistanceGA_Loja")
#Orion Directory to write .asc files and results
write.dir <-"./allcomb_151220/"

# setting up test
write.dir <-"./allcomb_151220_test/"



```

## Preparing raster surfaces

```{r}

# Read raster surfaces
ra1<- raster("./dem.loja_new.tif")
ra2 <- raster("./land.loja_new.tif")
ra3<- raster("./roads.loja_new.tif")

# stacks rasters
cat.stack <- stack(ra1, ra2, ra3)

# Change resolution to 250 m. Higher resolutions always cause the run to stop.
res <- c(250,250)

# Add CRS
crs <- "+proj=utm +zone=17 +south +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"

# dem 250 m res
dem.lowres <- projectRaster(cat.stack[[1]], res = res, crs = crs)

# land 250 m res
land.lowres <- projectRaster(cat.stack[[2]], crs = crs, res = res, method = "ngb")

# roads 250 m res
roads.lowres <- projectRaster(cat.stack[[3]], crs = crs, res = res, method = "ngb")


# Stack raster with 250 m resolution
lowres.stack <- stack(dem.lowres, land.lowres, roads.lowres)



# Reclassify categorical and feature rasters

# Reclassify land cover in a scale 1-100 based on an initial hypothesis of resistance. E.g. fragmented areas (cultivated/cropland) have lower values of resistance to Rhodnius ecuadoriensis.

#     1 Tree Cover, broadleaved, evergreen                        (8) 70               
#     2 Tree Cover, broadleaved, deciduous, closed                (7) 60
#     7 Tree Cover, regularly flooded, fresh  water (& brackish)  (10)90
#     12 Shrub Cover, closed-open, deciduous                      (4) 30
#     13 Herbaceous Cover, closed-open                            (5) 40
#     14 Sparse Herbaceous or sparse Shrub Cover                  (6) 50
#     15 Regularly flooded Shrub and/or Herbaceous Cover          (9) 80
#     16 Cultivated and managed areas                             (1) 1
#     17 Mosaic: Cropland / Tree Cover / Other natural vegetation (3) 20
#     18 Mosaic: Cropland / Shrub or Grass Cover                  (2) 10
#     20 Water Bodies (natural & artificial)                      (11) 100

land.reclass <- reclassify(lowres.stack[[2]], c(0, 1, 8,
                                             1.1, 2, 7,
                                             2.1, 7, 10,
                                             7.1, 12, 4,
                                             12.1, 13, 5,
                                             13.1, 14, 6,
                                             14.1, 15, 9,
                                             15.1, 16, 1,
                                             16.1, 17, 3,
                                             17.1, 18, 2,
                                             18.1, 20, 11))


# Reclassify roads
# Values 0-3 are roads #
# 0=Trails,           (2) 25
# 1=Highways,         (1) 1
# 2=secondary roads,  (3) 50
# 3=third roads       (4) 75
# 4=NA                (5) 100
#  1, 25, 50, 75, 100
roads.reclass <- reclassify(lowres.stack[[3]], c(-Inf, 0, 2,
                                              0.1, 1, 1,
                                              1.1, 2, 3,
                                              2.1, 3, 4,
                                              3.1,4, 5))




# Crop rasters to cover only data points to speed up optimization
ext <- extent(582149,688165,9500533,9560456) # set the new extent

dem.crop <- crop(lowres.stack[[1]], ext)
land.crop <- resample(land.reclass, dem.crop, method="ngb")
roads.crop <- resample(roads.reclass, dem.crop, method="ngb")

# Stack surfaces to optimize
loja.stack <- stack(dem.crop, land.crop, roads.crop)


```

## Preparing genetic matrix


```{r}

# Read in coordinate points
XY <- read.csv("./UTMpops.csv",header=TRUE)
# Create a site object that is a spatial points object with XY coordinates
sites.loja <- SpatialPoints(XY[ ,c(1,2)])

# Read in genetic distance matrix
genmat <- read.csv("./Pairwise_Gst_pops.csv", header = TRUE, row.names = 1)
genmat <- as.matrix(genmat)
Dgen <- as.dist(genmat)



```

## All combination Multisurface surface optimisation

All combinations of elevation, land cover and roads surfaces.

```{r, echo=FALSE}

### First test

GA.inputs <- GA.prep(ASCII.dir = loja.stack,
                     Results.dir = "all_comb",
                     select.trans = list("A", NA, NA),
                     pop.mult = 20,
                     pmutation = 0.2,
                     maxiter = 2,
                     parallel = 10)

gdist.inputs <- gdist.prep(length(sites.loja),
                           response = as.vector(Dgen),
                           samples = sites.loja,
                           method = 'commuteDistance') # Optimize using 


# Run all combinations

ac_multi <- all_comb(gdist.inputs = gdist.inputs, 
                     GA.inputs = GA.inputs,
                     results.dir = write.dir,
                     replicate = 3)

# second test


GA.inputs <- GA.prep(ASCII.dir = loja.stack,
                     Results.dir = "all_comb",
                     max.cat = 800,
                     max.cont = 800,
                     select.trans = list("A", NA, NA),
                     method = "AIC",
                     k.value = 3,
                     pop.mult = 20,
                     pmutation = 0.2,
                     maxiter = 2,
                     parallel = 10)

gdist.inputs <- gdist.prep(length(sites.loja),
                           response = as.vector(Dgen),
                           samples = sites.loja,
                           method = 'commuteDistance') # Optimize using 


# Run all combinations

ac_multi <- all_comb(gdist.inputs = gdist.inputs, 
                     GA.inputs = GA.inputs,
                     results.dir = write.dir,
                     replicate = 1)



### First complete run with no maximum number of iterations

# Orion directory
setwd("/shared2/salmosim/luis2/ecuadoriensis/landgen_analysis/ResistanceGA_Loja")
#Orion Directory to write .asc files and results

## this is the folder with the final results PLOS GENETICS
write.dir <-"./allcomb_151220/"



GA.inputs <- GA.prep(ASCII.dir = loja.stack,
                     Results.dir = "all_comb",
                     select.trans = list("M", NA, NA),
                     pop.mult = 20,
                     parallel = 10)

gdist.inputs <- gdist.prep(length(sites.loja),
                           response = as.vector(Dgen),
                           samples = sites.loja,
                           method = 'commuteDistance')


# Run all combinations

ac_multi <- all_comb(gdist.inputs = gdist.inputs, 
                     GA.inputs = GA.inputs,
                     results.dir = write.dir,
                     replicate = 3)





